---
title: "NBA Alt Awards 2025 (hoopR & NBA PBP)"
author: "Sumitro Datta"
format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{typearea}
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
packages=c("tidyverse",
           "hoopR",
           "janitor", # cleaning variable names
           "teamcolors", # team color palettes
           "ggdark", # dark background plots
           "gt") # table creator
for (pkg in packages){
  if(!require(pkg,character.only = TRUE)){
    install.packages(pkg,repos = "http://cran.us.r-project.org")
  }
}
```

```{r full_pbp_load}
current_year=2025
current_year_for_pbp=paste0(current_year-1,"-",str_sub(current_year,-2))
pbp_curr_yr=read_csv(paste0("Data/",current_year_for_pbp,"_pbp.csv")) %>% 
  group_by(GAME_ID) %>% 
  fill(c(SCORE,SCOREMARGIN)) %>%
  replace_na(list(SCORE="0 - 0",SCOREMARGIN="TIE")) %>% ungroup() %>% 
  mutate(SCOREMARGIN=if_else(SCOREMARGIN=="TIE",0,as.numeric(SCOREMARGIN))) %>%
  separate(SCORE,sep=" - ",into=c("visitor_score","home_score"),convert = TRUE) %>%
  mutate(clutch=(PERIOD==4 & abs(SCOREMARGIN)<=5) & PCTIMESTRING<= hms::hms(hours=5)|(PERIOD>4)) %>% ungroup()
```

# Awards

## The Spark Plug Award (sponsored by Lt. Surge, presented by American Express CEO Stephen J Squeri)

Most charges drawn per 36 minutes (minimum 70% of games played), credit to morron88 for the idea to separate charges & loose balls in 2020

```{r hustle_full}
compact_standings=nba_leaguestandingsv3()$Standings %>% 
  type_convert() %>%
  select(TeamID,TeamCity,TeamName,Conference,Division,WINS,LOSSES) %>% 
  mutate(TeamGP=WINS+LOSSES,TeamFullName=paste(TeamCity,TeamName))

player_bio=nba_leaguedashplayerbiostats()$LeagueDashPlayerBioStats %>% 
  select(PLAYER_ID:GP) %>%
  type_convert()

team_info=left_join(compact_standings,
          player_bio %>% 
            group_by(TEAM_ID) %>% 
            slice(1) %>% 
            select(TEAM_ID,TEAM_ABBREVIATION), by=join_by(TeamID==TEAM_ID))

player_w_gp_percentages=left_join(
  player_bio,
  compact_standings %>% select(contains("Team")),
  by=join_by(TEAM_ID==TeamID)
  ) %>%
  mutate(G_PERCENT=GP/TeamGP)

hustle=nba_leaguehustlestatsplayer(per_mode="PerGame")$HustleStatsPlayer %>% type_convert()
hustle_w_gp_qualify=left_join(hustle,
                              player_w_gp_percentages %>% 
                                select(PLAYER_ID,G_PERCENT))
```

```{r charges_drawn}
hustle_70_percent_gp=hustle_w_gp_qualify %>% 
  filter(G_PERCENT>=0.7) %>%
  mutate(across(c(CHARGES_DRAWN,
                  LOOSE_BALLS_RECOVERED,
                  DEFLECTIONS,
                  SCREEN_AST_PTS),~./MIN*36))

hustle_70_percent_gp %>% slice_max(CHARGES_DRAWN,n=10) %>% 
  select(PLAYER_NAME,CHARGES_DRAWN) %>% gt() %>%
  fmt_number(-c(PLAYER_NAME),decimals=4) %>%
  as_raw_html()
```

## The Most Loose Balls Recovered Award (sponsored by Hungry Hungry Hippos, presented by Dennis Rodman & [Nene's doctor](https://www.espn.com/nba/news/story?id=3197423))

Per 36 minutes, minimum 70% of games played

```{r loose_balls}
hustle_70_percent_gp %>% slice_max(LOOSE_BALLS_RECOVERED,n=10) %>% 
  select(PLAYER_NAME,LOOSE_BALLS_RECOVERED) %>% gt() %>%
  fmt_number(-c(PLAYER_NAME),decimals=4)%>%
  as_raw_html()
```

## The Plexiglass Award

most deflections per 36 minutes, minimum 70% of games played

```{r deflections}
hustle_70_percent_gp %>% slice_max(DEFLECTIONS,n=10) %>% 
  select(PLAYER_NAME,DEFLECTIONS) %>% gt() %>%
  fmt_number(-c(PLAYER_NAME),decimals=4) %>%
  as_raw_html()
```

## The Wes Unseld Memorial Brick Wall Award

most points generated by screen assists per 36 minutes, minimum 70% of games played

```{r screen_ast_pts}
hustle_70_percent_gp %>% slice_max(SCREEN_AST_PTS,n=10) %>% 
  select(PLAYER_NAME,SCREEN_AST_PTS) %>% gt() %>%
  fmt_number(-c(PLAYER_NAME),decimals=4) %>%
  as_raw_html()
```

\newpage

\KOMAoptions{paper=landscape,pagesize}

## The "He Trick Y'All, Running Around, Doing Nothing" Award (sponsored by Russell Westbrook, presented by Tony Snell)

Lowest sum of per-36 percentile ranks in the following: charges, contested shots, deflections, defensive boxouts, defensive loose balls recovered (minimum 50% of games played)

```{r hustle_ranks_bottom}
hustle_50_percent_gp=hustle_w_gp_qualify %>% filter(G_PERCENT>=0.5) %>% 
  select(PLAYER_NAME,MIN,CHARGES_DRAWN, CONTESTED_SHOTS_2PT,CONTESTED_SHOTS_3PT, 
         DEFLECTIONS,DEF_BOXOUTS,DEF_LOOSE_BALLS_RECOVERED) %>%
  mutate(across(-c(PLAYER_NAME,MIN),~percent_rank(./MIN*36))) %>%
  mutate(sum=rowSums(across(CHARGES_DRAWN:DEF_LOOSE_BALLS_RECOVERED)))

write_csv(hustle_50_percent_gp,"Output Data/Hustle Ranks.csv")

hustle_50_percent_gp %>% slice_min(sum,n=10) %>% gt() %>% 
  fmt_percent(columns=-c(PLAYER_NAME,MIN)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(-c(PLAYER_NAME,MIN,sum)~px(115))
```

```{r westbrook_hustle_rank}
hustle_50_percent_gp %>% 
  filter(str_detect(PLAYER_NAME,"Russell Westbrook")) %>% 
  gt() %>% 
  fmt_percent(columns=-c(PLAYER_NAME,MIN)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(-c(PLAYER_NAME,MIN,sum)~px(115))
```

\newpage

## The "Got that Dawg in Him" Award (presented by Air Bud)\*

Highest sum of per-36 percentile ranks in the following: charges, contested shots, deflections, defensive boxouts, defensive loose balls recovered (minimum 50% of games played) (credit to memeticengineering for the idea)

```{r hustle_ranks_top}
hustle_50_percent_gp %>% slice_max(sum,n=10) %>% gt() %>% 
  fmt_percent(columns=-c(PLAYER_NAME,MIN)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(-c(PLAYER_NAME,MIN,sum)~px(115))
```

## The Trickshot Grenadier Award (presented by Dude Perfect)

Highest sum of percentile ranks in FGA, FGA frequency & eFG% on shots with 4 seconds or less on the shotclock (credit to BehavioralSink & Bylanta for the idea)

```{r trickshot_grenadier}
late_shotclock_shots=nba_leaguedashplayerptshot(shot_clock_range="4-0 Very Late")$LeagueDashPTShots %>% type_convert() %>% select(PLAYER_ID:EFG_PCT)

trickshot_grenades=late_shotclock_shots %>% 
mutate(across(c(FGA_FREQUENCY,FGA,EFG_PCT),
              percent_rank,
              .names="{.col}_percentile")) %>%
  mutate(sum_percentiles=rowSums(across(contains("percentile"))))

write_csv(trickshot_grenades,"Output Data/Trickshot Grenadier.csv")

trickshot_grenades %>%
  slice_max(sum_percentiles,n=10) %>%
  select(PLAYER_NAME,FGA_FREQUENCY,FGA,EFG_PCT,contains("percentile")) %>%
  gt() %>%
  fmt_percent(columns=-c(PLAYER_NAME,FGA)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(contains("FREQ")~px(105),
             -c(PLAYER_NAME,FGA)~px(100))
```

\newpage

\KOMAoptions{paper=portrait,pagesize}

## The Clutch Line Syndrome Award (presented by Diar DeRozan)

biggest decline from non-clutch FT% to clutch FT% (credit to Necessary_Career_253 for the idea)

```{r clutch_ft_worst}
free_throws=pbp_curr_yr %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Free Throw")))

clutch_free_throws=free_throws %>% filter(clutch) %>% 
  group_by(PLAYER1_NAME) %>%
  summarize(clutch_fts_attempted=n(),
            clutch_fts_missed=sum(str_detect(VISITORDESCRIPTION,"MISS")|
                                    str_detect(HOMEDESCRIPTION,"MISS"),
                                  na.rm=TRUE),
            clutch_ft_percent=1-clutch_fts_missed/clutch_fts_attempted) %>% 
  ungroup()

non_clutch_free_throws=free_throws %>% filter(!clutch) %>% 
  group_by(PLAYER1_NAME) %>%
  summarize(non_clutch_fts_attempted=n(),
            non_clutch_fts_missed=sum(str_detect(VISITORDESCRIPTION,"MISS")|
                                        str_detect(HOMEDESCRIPTION,"MISS"),
                                      na.rm=TRUE),
            non_clutch_ft_percent=1-non_clutch_fts_missed/non_clutch_fts_attempted) %>% ungroup()

clutch_differential=inner_join(clutch_free_throws,non_clutch_free_throws) %>% 
  mutate(differential=clutch_ft_percent-non_clutch_ft_percent) %>% 
  filter(clutch_fts_attempted>=10)

write_csv(clutch_differential,"Output Data/Clutch Free Throw Differential.csv")

clutch_differential %>% slice_min(differential,n=10) %>% 
  select(PLAYER1_NAME,clutch_fts_attempted,
         clutch_ft_percent,non_clutch_ft_percent,differential) %>%
  gt() %>% fmt_percent(columns=-c(PLAYER1_NAME,clutch_fts_attempted)) %>%
  as_raw_html()
```

## The "Ice, Ice, Baby" Award (sponsored by Hisense)

biggest improvement from non-clutch FT% to clutch FT%

```{r clutch_ft_best}
clutch_differential %>% slice_max(differential,n=10) %>% 
  select(PLAYER1_NAME,clutch_fts_attempted,
         clutch_ft_percent,non_clutch_ft_percent,differential) %>%
  gt() %>% fmt_percent(columns=-c(PLAYER1_NAME,clutch_fts_attempted)) %>%
  as_raw_html()
```

## The Dikembe Mutombo Memorial "No Fly Zone" Award\*

most blocked dunks as the blocking player

```{r no_fly_zone}
misses=pbp_curr_yr %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_starts(.,"MISS")))
missed_dunks=misses %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Dunk")))
blocked_dunks=missed_dunks %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"BLOCK")))
#as the blocker
blocked_dunks %>% group_by(PLAYER3_NAME) %>% 
  summarize(count=n()) %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The Rejected for Boarding Award (sponsored by United Airlines)\*

most blocked dunks as the dunking player (credit to Legdrop_soup for the idea and asw7412 for the sponsor)

```{r reject_boarding}
#as the blockee
blocked_dunks %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The No Time to Taunt Award (presented by Tim Duncan)\*

highest percent of blocks that stayed inbounds & recovered by blocker's team (credit to gibberisle for the idea)

```{r no_time_to_taunt}

pbp_curr_yr_w_next_play=pbp_curr_yr %>% 
  group_by(GAME_ID) %>%
  mutate(next_home_play=lead(HOMEDESCRIPTION),
         next_visitor_play=lead(VISITORDESCRIPTION)) %>% ungroup()

blocks=pbp_curr_yr_w_next_play %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"BLOCK"))) %>%
  mutate(home_or_visitor_block=if_else(
    str_detect(HOMEDESCRIPTION,"BLOCK"),"home","visitor"))

team_rebounds=pbp_curr_yr_w_next_play %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_ends(.,"Rebound")))

blocks_wo_team_rebs=anti_join(blocks,team_rebounds,by=c("GAME_ID","PERIOD","PCTIMESTRING")) %>% 
  mutate(recovered_by_which_team=case_when(
    home_or_visitor_block=="home" & is.na(next_visitor_play)~"own",
    home_or_visitor_block=="visitor" & is.na(next_home_play)~"own",
    TRUE~"other"))

player_blocks=blocks %>% group_by(PLAYER3_NAME) %>% summarize(blocks=n())

player_blocks_wo_team_rebs=blocks_wo_team_rebs %>% 
  group_by(PLAYER3_NAME,recovered_by_which_team) %>% summarize(n=n()) %>%
  pivot_wider(names_from=recovered_by_which_team,values_from=n,values_fill=0)

player_blk_info=left_join(player_blocks,player_blocks_wo_team_rebs) %>%
  mutate(percent_blk_recovered_by_own=own/blocks)

average_games_played=compact_standings %>% 
  summarize(average_GP=mean(TeamGP)) %>% pull(average_GP)

player_blk_info %>% filter(blocks>=average_games_played*0.5) %>% 
  slice_max(percent_blk_recovered_by_own,n=10) %>% gt() %>%
  fmt_percent(columns=percent_blk_recovered_by_own) %>%
  as_raw_html()
```

## The "David vs Goliath" Award (presented by Dwyane Wade)

most shots blocked where the blocker is at least 5 inches shorter than the blockee

```{r david_v_goliath}
player_bio=nba_leaguedashplayerbiostats()$LeagueDashPlayerBioStats %>% type_convert()

block_player_heights=left_join(blocks,
          player_bio %>% select(PLAYER_ID,
                                blocked_player_height=PLAYER_HEIGHT_INCHES),
          by=join_by(PLAYER1_ID==PLAYER_ID)) %>%
  left_join(.,player_bio %>% select(PLAYER_ID,
                                blocking_player_height=PLAYER_HEIGHT_INCHES),
            by=join_by(PLAYER3_ID==PLAYER_ID)) %>% 
  mutate(height_diff=blocked_player_height-blocking_player_height)

block_player_heights %>% filter(height_diff>=5) %>%
  group_by(PLAYER3_NAME) %>% summarize(count=n()) %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The "FUCK OUTTA HERE, I GOT THAT SHIT" Award (presented by Carmelo Anthony)

Lowest contested rebound percentage, minimum 50% of games played

```{r rebounding}
rebounding=nba_playerdashptreb(team_id=0,player_id=0)$OverallRebounding %>%
  type_convert()

reb_w_gp_qualify=left_join(
  rebounding,
  player_w_gp_percentages %>% select(PLAYER_ID,G_PERCENT,PLAYER_HEIGHT_INCHES)
  ) %>%
  filter(G_PERCENT>=0.5)
```

```{r melo_rebound}
reb_w_gp_qualify %>% slice_min(C_REB_PCT,n=10) %>% 
  select(PLAYER_NAME_LAST_FIRST,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

alternatively: restricting to players \> 6 foot 6 inches in height

```{r rebound_above_66}
reb_w_gp_qualify %>% filter(PLAYER_HEIGHT_INCHES>6*12+6) %>% 
  slice_min(C_REB_PCT,n=10) %>% 
  select(PLAYER_NAME_LAST_FIRST,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

## The "Glass Cleaner" Award (presented by Dennis Rodman, sponsored by Windex)

Highest contested rebound percentage, minimum 50% of games played

```{r glass_cleaner}
reb_w_gp_qualify %>% slice_max(C_REB_PCT,n=10) %>% 
  select(PLAYER_NAME_LAST_FIRST,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

alternatively: restricting to players \< 6 foot 7 inches in height

```{r rebound_below_67}
reb_w_gp_qualify %>% filter(PLAYER_HEIGHT_INCHES<6*12+7) %>% 
  slice_max(C_REB_PCT,n=10) %>% 
  select(PLAYER_NAME_LAST_FIRST,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

## The "Call Game" Award (presented by Paul "No OT Tonight" George)

most game winning points (defined as the first points that eclipsed the losing team's total) (credit to Necessary_Career_253 for the idea & Clownp3nis for the presenter)

```{r game_winners}
game_results=pbp_curr_yr %>% group_by(GAME_ID) %>% 
  summarize(vis_final=max(visitor_score),home_final=max(home_score)) %>% 
  ungroup() %>%
  mutate(winner=if_else(home_final>vis_final,"home","visitor"),
         loser_points=if_else(home_final>vis_final,vis_final,home_final))
game_winning_shots=left_join(pbp_curr_yr,game_results) %>%
  filter((winner=="home" & home_score>loser_points)|
           (winner=="visitor" & visitor_score>loser_points)) %>%
  group_by(GAME_ID) %>% slice_min(EVENTNUM) %>% ungroup() %>%
  mutate(shot_value=case_when(
    str_detect(HOMEDESCRIPTION,"Free Throw")~1,
    str_detect(VISITORDESCRIPTION,"Free Throw")~1,
    str_detect(VISITORDESCRIPTION,"3PT")~3,
    str_detect(HOMEDESCRIPTION,"3PT")~3,
    TRUE~2
  )) %>%
  group_by(PLAYER1_NAME) %>% 
  summarize(game_winning_3=sum(shot_value==3),
            game_winning_2=sum(shot_value==2),
            game_winning_ft=sum(shot_value==1),
            game_winning_points=sum(shot_value))

write_csv(game_winning_shots,"Output Data/Game-Winning Shots.csv")

game_winning_shots %>% slice_max(game_winning_points,n=5) %>%
  gt() %>%
  as_raw_html()
```

## The "Fix My Own Mistakes" Award

most shots after rebounding own missed shot (credit to Necessary_Career_253 for the idea)

```{r fix_own_mistake}
pbp_curr_yr_w_before_after=pbp_curr_yr %>% 
  group_by(GAME_ID) %>%
  mutate(next_home_play=lead(HOMEDESCRIPTION),
         next_visitor_play=lead(VISITORDESCRIPTION),
         next_player=lead(PLAYER1_ID),
         last_home_play=lag(HOMEDESCRIPTION),
         last_visitor_play=lag(VISITORDESCRIPTION),
         last_player=lag(PLAYER1_ID)) %>% ungroup()

shot_after_own_rebound_miss=pbp_curr_yr_w_before_after %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"REBOUND"))) %>%
  filter(PLAYER1_ID==next_player,PLAYER1_ID==last_player) %>%
  filter(str_detect(next_home_play,"PTS")|
           str_detect(next_visitor_play,"PTS")) %>% 
  mutate(shot_value=case_when(
    str_detect(next_home_play,"3PT")~3,
    str_detect(next_visitor_play,"3PT")~3,
    TRUE~2
  )) %>% 
  group_by(PLAYER1_NAME) %>% 
  summarize(fix_with_3=sum(shot_value==3),
            fix_with_2=sum(shot_value==2),
            fixed_shots=n(),
            fixed_points=sum(shot_value))

shot_after_own_rebound_miss %>% slice_max(fixed_points,n=10) %>%
  gt() %>%
  as_raw_html()
```

## The "Oops, I Dunked It Again" Award (sponsored by Britney Spears, presented by Gary Payton & Shawn Kemp)\*

Most prolific alley-oop duo, both ways (credit to lactardenthusiast for the idea)

```{r alley_oops}
#get alley oops
alley_oops=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Alley")))
#remove missed alley oop attempts
made_alley_oops=alley_oops %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_starts(.,"MISS",negate=TRUE)))
#remove alley oops that have missing 2nd player (self-alley oop or perhaps missing data?)
made_alley_oops_complete=made_alley_oops %>% 
  filter(!is.na(PLAYER2_NAME)) %>% rowwise() %>% 
  mutate(p1=min(PLAYER1_NAME,PLAYER2_NAME),p2=max(PLAYER1_NAME,PLAYER2_NAME)) %>%
  ungroup()

made_alley_oops_complete %>% group_by(p1,p2) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The "Human Highlight Reel" Award (presented by Dominique Wilkins)\*

players that had the most games with 2+ dunks (credit to frosiano for the idea)

```{r two_plus_dunks}
dunks=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Dunk")))
missed_dunks=dunks %>%
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"MISS")))
made_dunks=anti_join(dunks,missed_dunks)

made_dunks %>% group_by(GAME_ID,PLAYER1_NAME) %>% summarize(dunks_in_game=n()) %>%
  filter(dunks_in_game>=2) %>% ungroup() %>% group_by(PLAYER1_NAME) %>% 
  summarize(games_w_2plus_dunks=n()) %>% slice_max(games_w_2plus_dunks,n=10) %>% gt() %>%
  as_raw_html()
```

## The "Just the Tip" Award

best jump ball win percentage, minimum 5 jump balls (credit to livejamie)

```{r jump_balls}
jump_balls=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Jump Ball")))

jump_balls_by_player=jump_balls %>% 
  select(PLAYER1_ID:PLAYER3_TEAM_ABBREVIATION) %>%
  mutate(winner=if_else(PLAYER1_TEAM_ABBREVIATION==PLAYER3_TEAM_ABBREVIATION,
                        PLAYER1_NAME,PLAYER2_NAME),
         loser=if_else(winner==PLAYER1_NAME,
                       PLAYER2_NAME,PLAYER1_NAME)) %>%
  select(winner,loser) %>% 
  pivot_longer(winner:loser,values_to = "player") %>% 
  group_by(player) %>%
  summarize(jump_balls_participated=n(),
            wins=sum(name=="winner"),
            losses=sum(name=="loser"),
            win_percent=wins/jump_balls_participated)

jump_ball_percentiles=jump_balls_by_player %>% filter(jump_balls_participated>=5) %>%
  mutate(across(c(jump_balls_participated,win_percent),
                percent_rank,
                .names="{.col}_percentile")) %>%
  mutate(avg_percentiles=rowMeans(across(contains("percentile"))))

write_csv(jump_ball_percentiles,"Output Data/Jump Ball Percentiles.csv")

jump_ball_percentiles %>%
  slice_max(avg_percentiles,n=10) %>%
  select(player,jump_balls_participated,win_percent,contains("percentile")) %>%
  gt() %>% fmt_percent(columns=-c(player:jump_balls_participated)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(-player~px(115))
```

## The Bowling Ball Award (sponsored by [Pete Weber](#0), presented by Glen "Big Baby" Davis)\*

most charges committed (credit to Kdog122025 for the idea)

```{r bowling_ball}
charges=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Charge")))

charges %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## "The Good Ol' Hockey Game, is the Best Game You Can Name" Award (presented by Dominik Hasek)\*

most goaltends committed (credit to Kdog122025 for the idea)

```{r goaltends}
goaltends=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Goaltending")))

goaltends %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## "The Thing about Arsenal Is They Always Try to Walk It In" Award (presented by MLS Commissioner Don Garber)\*

most kicked ball violations

```{r kicked_balls}
kicked_balls=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Kick")))

kicked_balls %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The "Pack Your Bags" Award (Sponsored by Emirates, the global airline partner of the NBA)

Player with the most traveling calls

```{r travelling}
traveling=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Traveling")))

traveling %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The "Counting to 3" Award (Sponsored by the National Parks System, presented by Sesame Street's Count von Count)

Player with the most illegal defense/defensive 3-second calls (credit to PsychoM & MrBuckBuck for the idea)

```{r def_3_seconds}
def_3_seconds=pbp_curr_yr %>% 
    filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"3 Second")))

def_3_seconds %>% group_by(PLAYER1_NAME) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The Zaza Pachulia All-Ball Award\*

most 3-pt shooting fouls committed (credit to watchingsongsDL, kingcobweb & An-Indian-In-The-NBA for the idea, and sunnysideoutside for the name)

```{r 3_pt_fouls_commit}

third_free_throws=pbp_curr_yr %>%
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Free Throw 3 of 3")))

shooting_fouls=pbp_curr_yr %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"S.FOUL")))

fouls_on_missed_threes=inner_join(shooting_fouls,third_free_throws,
                                  by=c('GAME_ID','PERIOD','PCTIMESTRING'))

three_pointers=pbp_curr_yr %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"3PT")))

made_threes=three_pointers %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_starts(.,"MISS",negate=TRUE)))

fouls_on_made_threes=inner_join(shooting_fouls,made_threes,
                                  by=c('GAME_ID','PERIOD','PCTIMESTRING'))

fouls_on_threes=fouls_on_missed_threes %>% add_row(fouls_on_made_threes)

fouls_on_threes %>% group_by(PLAYER1_NAME.x) %>% 
  summarize(count=n()) %>% ungroup() %>% slice_max(count,n=10) %>% gt() %>%
  as_raw_html()
```

## The "I'll Have It to Go" Award (sponsored by DoorDash)

coach with lowest timeout utilization (credit to xfinityhomeboy, Ill_Ad3517 & s-sea (who also came up with the name))

```{r most_timeouts_left_over}
num_timeouts_available=pbp_curr_yr %>% 
  filter(!is.na(PLAYER1_TEAM_ABBREVIATION)) %>% 
  group_by(GAME_ID,PERIOD,PLAYER1_TEAM_ABBREVIATION) %>% slice(1) %>% 
  select(GAME_ID,PERIOD,PLAYER1_TEAM_ABBREVIATION) %>% ungroup() %>%
  group_by(team=PLAYER1_TEAM_ABBREVIATION) %>% 
  summarize(games=sum(PERIOD<=4)/4,
            overtimes=sum(PERIOD>4),
            num_timeouts=7*games+2*overtimes) %>% 
  left_join(.,team_info %>% 
              select(TeamName,TEAM_ABBREVIATION),
            by=join_by(team==TEAM_ABBREVIATION))

num_timeouts_used=pbp_curr_yr %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Timeout"))) %>% 
  filter(if_any(c(HOMEDESCRIPTION,VISITORDESCRIPTION),
                ~str_detect(.,"Excess",negate=TRUE))) %>%
  select(GAME_ID,PERIOD,HOMEDESCRIPTION,VISITORDESCRIPTION) %>%
  pivot_longer(cols=HOMEDESCRIPTION:VISITORDESCRIPTION,
               values_to="timeout") %>%
  select(-name) %>% filter(!is.na(timeout)) %>%
  mutate(team_name=str_to_title(word(timeout,1,sep=" Timeout:"))) %>%
  group_by(team_name) %>% summarise(num_timeouts_used=n())

timeouts_used_percent=left_join(num_timeouts_available,
                                num_timeouts_used,
                                by=join_by(TeamName==team_name)) %>% 
  mutate(timeout_use_percent=num_timeouts_used/num_timeouts) %>% 
  arrange(desc(timeout_use_percent))

timeouts_used_percent %>% 
  select(team,num_timeouts,
         num_timeouts_used,timeout_use_percent) %>%
  slice_min(timeout_use_percent,n=5) %>%
  gt() %>%
  fmt_percent(columns=timeout_use_percent) %>%
  as_raw_html()
```

## The "Fine, I'll Do It Myself" Award (sponsored by Thanos, presented by Allen Iverson)

Highest percentage of unassisted field goals, minimum 50% of games played

```{r field_goals_ast_and_uast,eval=FALSE}
assisted_fg_w_gp_qualify=tibble()
player_ids_over_50_percent_gp=player_w_gp_percentages %>% 
  filter(G_PERCENT>=0.5) %>% pull(PLAYER_ID)
for (player_id in player_ids_over_50_percent_gp){
  new_player_shooting_splits=nba_playerdashboardbyshootingsplits(
    player_id = as.character(player_id)
    )$OverallPlayerDashboard %>%
  select(FGM,FGA,PCT_AST_FGM,PCT_UAST_FGM) %>% mutate(PLAYER_ID=player_id)
  assisted_fg_w_gp_qualify=bind_rows(
    assisted_fg_w_gp_qualify,
    new_player_shooting_splits)
}

assisted_fg_w_gp_qualify_player_names=left_join(
  assisted_fg_w_gp_qualify,
  player_bio %>% select(PLAYER_ID:PLAYER_NAME))
```

```{r do_it_myself,eval=FALSE}
assisted_fg_w_gp_qualify_player_names %>%
  slice_max(PCT_UAST_FGM,n=10) %>%
  select(PLAYER_NAME,FGA,PCT_UAST_FGM) %>% gt() %>%
  as_raw_html()
```

## The "You Gotta Feed Me" Award (presented by Joey Chestnut & Marcin Gortat)

Highest percentage of assisted field goals, minimum 50% of games played

```{r gotta_feed,eval=FALSE}
assisted_fg_w_gp_qualify_player_names %>%
  slice_max(PCT_AST_FGM,n=10) %>%
  select(PLAYER_NAME,FGA,PCT_AST_FGM) %>% gt() %>%
  as_raw_html()
```
