---
title: "NBA Alt Awards 2026 (Scraping)"
author: "Sumitro Datta"
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","rvest","polite","rJava","tabulapdf","RKaggle",
               "janitor", "gt")
```

```{r data_load}
current_year=2026

play_by_play=read_csv("Data/Player Play by Play.csv") %>% 
  filter(season==current_year)

team_info=read_csv("Data/Team Summaries.csv") %>% 
  filter(season==current_year,!is.na(abbreviation)) %>% 
  select(team,abbreviation)
teams=team_info %>% pull(abbreviation)

western_conf=c("DAL","DEN","GSW","HOU","LAC","LAL","MEM",
               "MIN","NOP","OKC","PHO","POR","SAC","SAS",
               "UTA")

teams_w_conf=tibble(tm=teams) %>% 
  mutate(conference=case_when(tm %in% western_conf~"West",
                              TRUE~"East"))

bbref_bow=bow("https://www.basketball-reference.com/",
              user_agent = "Sumitro Datta",force=TRUE,delay = 10)
```

# Awards

## The "Home Cooking" Award (sponsored by Julia Child)

biggest PPG increase from road to home, min 50% of games played and 10 PPG (credit to livejamie and psychoM for the idea)

```{r box_scores}
x2025_2026_box_scores=
  get_dataset("eoinamoore/historical-nba-data-and-player-box-scores") %>%
  .[[6]] %>% 
  filter(gameDateTimeEst>=mdy("10/21/2025"),gameType != "All-Star Game") %>% 
  filter(!is.na(numMinutes) & !is.na(win) & numMinutes>=0.01) %>% 
  mutate(game_score=
           points+fieldGoalsMade*0.4-0.7*fieldGoalsAttempted-
           0.4*(freeThrowsAttempted-freeThrowsMade)+
           0.7*reboundsOffensive+0.3*reboundsDefensive+steals+
           0.7*assists+0.7*blocks-0.4*foulsPersonal-turnovers)
```

```{r home_away_split_process}
home_away_ppg_gmsc=x2025_2026_box_scores %>%
  group_by(personId,firstName,lastName,home) %>% 
  summarize(n=n(),points=sum(points),game_score=sum(game_score)) %>%
  mutate(home=if_else(home==0,"away","home"),
         name=paste(firstName,lastName)) %>%
  pivot_wider(names_from=home,values_from = c(n,points,game_score),
              values_fill = 0) %>%
  mutate(g_total=n_home+n_away,
         pts_total=points_home+points_away,
         ppg_away=points_away/n_away,
         ppg_home=points_home/n_home,
         ppg_total=pts_total/g_total,
         ppg_diff=ppg_home-ppg_away,
         gmsc_away=game_score_away/n_away,
         gmsc_home=game_score_home/n_home,
         gmsc_diff=gmsc_home-gmsc_away) %>%
  ungroup()

write_csv(
  home_away_ppg_gmsc %>% 
  select(-c(firstName,lastName,points_away:game_score_home,pts_total)) %>%
  relocate(personId,name,n_away,ppg_away,n_home,ppg_home,ppg_diff,
           contains("gmsc"),contains("total")),
  "Output Data/Home-Away PPG GmSc Splits.csv")
```

```{r homebody_ppg}
#| tbl-pos: T
#| layout-ncol: 1

home_away_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2,ppg_total>=10) %>% 
  slice_max(ppg_diff,n=5) %>%
  select(name,n_away,ppg_away,n_home,ppg_home,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  tab_spanner(label="Home",columns=ends_with("home")) %>%
  tab_spanner(label="Away",columns=ends_with("away")) %>%
  cols_label(n_away="GP",ppg_away="PPG",n_home="GP",ppg_home="PPG")
```

Alternatively: biggest game score increase from road to home, min 50% of games played

```{r homebody_gmsc}
#| tbl-pos: T
#| layout-ncol: 1

home_away_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2) %>% slice_max(gmsc_diff,n=5) %>%
  select(name,n_away,gmsc_away,n_home,gmsc_home,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  tab_spanner(label="Home",columns=ends_with("home")) %>%
  tab_spanner(label="Away",columns=ends_with("away")) %>%
  cols_label(n_away="GP",gmsc_away="GmSc",n_home="GP",gmsc_home="GmSc")
```

\newpage

## The "Jetsetter" Award (presented by Eric Bledsoe)

biggest PPG decrease from road to home, min 50% of games played and 10 PPG (credit to livejamie and psychoM for the idea)

```{r jetsetter_ppg}
#| tbl-pos: T
#| layout-ncol: 1

home_away_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2,ppg_total>=10) %>% 
  slice_min(ppg_diff,n=5) %>%
  select(name,n_away,ppg_away,n_home,ppg_home,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  tab_spanner(label="Home",columns=ends_with("home")) %>%
  tab_spanner(label="Away",columns=ends_with("away")) %>%
  cols_label(n_away="GP",ppg_away="PPG",n_home="GP",ppg_home="PPG")
```

Alternatively: biggest game score decrease from road to home, min 50% of games played

```{r jetsetter_gmsc}
#| tbl-pos: T
#| layout-ncol: 1

home_away_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2) %>% 
  slice_min(gmsc_diff,n=5) %>%
  select(name,n_away,gmsc_away,n_home,gmsc_home,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  tab_spanner(label="Home",columns=ends_with("home")) %>%
  tab_spanner(label="Away",columns=ends_with("away")) %>%
  cols_label(n_away="GP",gmsc_away="GmSc",n_home="GP",gmsc_home="GmSc")
```

## The Bright Lights Merchant Award

biggest PPG increase from playing against teams below 45% win percentage to playing against teams above 55% win percentage, min 50% of games played and 10 PPG

```{r opp_win_percent_split_process}
standings=x2025_2026_box_scores %>% 
  #remove nba cup final
  filter(gameId != 62500001) %>% 
  group_by(gameId,playerteamName) %>%
  reframe(tot_points=sum(points)) %>% group_by(gameId) %>%
  mutate(winner=max(tot_points)==tot_points) %>% ungroup() %>%
  group_by(playerteamName) %>% reframe(games=n(),wins=sum(winner)) %>%
  mutate(losses=games-wins,win_percent=wins/games)

opp_win_percent_ppg_gmsc=left_join(
  x2025_2026_box_scores,
  standings %>% 
    mutate(category=case_when(win_percent>=0.55~"above_55_w_percent",
                              win_percent<=0.45~"below_45_w_percent",
                              TRUE~"middle")) %>%
    select(team=playerteamName,category),
  by=join_by(opponentteamName==team)
) %>%
  group_by(personId,firstName,lastName,category) %>% 
  summarize(n=n(),points=sum(points),game_score=sum(game_score)) %>%
  mutate(name=paste(firstName,lastName)) %>%
  pivot_wider(names_from=category,values_from = c(n,points,game_score),
              values_fill = 0) %>%
  mutate(g_total=n_above_55_w_percent+n_below_45_w_percent+n_middle,
         pts_total=points_above_55_w_percent+
           points_below_45_w_percent+
           points_middle,
         ppg_below_45_w_percent=
           points_below_45_w_percent/n_below_45_w_percent,
         ppg_above_55_w_percent=
           points_above_55_w_percent/n_above_55_w_percent,
         ppg_total=pts_total/g_total,
         ppg_diff=ppg_above_55_w_percent-ppg_below_45_w_percent,
         gmsc_below_45_w_percent=
           game_score_below_45_w_percent/n_below_45_w_percent,
         gmsc_above_55_w_percent=
           game_score_above_55_w_percent/n_above_55_w_percent,
         gmsc_diff=gmsc_above_55_w_percent-gmsc_below_45_w_percent) %>%
  ungroup()

write_csv(
  opp_win_percent_ppg_gmsc %>% 
  select(-c(firstName,lastName,n_middle,
            points_above_55_w_percent:game_score_middle,pts_total)) %>%
  relocate(personId,name,
           n_below_45_w_percent,ppg_below_45_w_percent,
           n_above_55_w_percent,ppg_above_55_w_percent,
           ppg_diff,
           contains("gmsc"),contains("total")),
  "Output Data/Opponent Strength PPG GmSc Splits.csv")
```

```{r bright_lights_merchant_ppg}
#| tbl-pos: T
#| layout-ncol: 1

opp_win_percent_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2,ppg_total>=10) %>% 
  slice_max(ppg_diff,n=5) %>%
  select(name,n_below_45_w_percent,
         ppg_below_45_w_percent,n_above_55_w_percent,
         ppg_above_55_w_percent,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  tab_spanner(label="<= 45% Win %",columns=ends_with("45_w_percent")) %>%
  tab_spanner(label=">= 55% Win %",columns=ends_with("55_w_percent")) %>%
  cols_label(
    n_below_45_w_percent="GP",ppg_below_45_w_percent="PPG",
    n_above_55_w_percent="GP",ppg_above_55_w_percent="PPG"
  )
```

\newpage

Alternatively: biggest game score increase from playing against teams below 45% win percentage to playing against teams above 55% win percentage, min 50% of games played

```{r bright_lights_merchant_gmsc}
#| tbl-pos: T
#| layout-ncol: 1

opp_win_percent_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2) %>% 
  slice_max(gmsc_diff,n=5) %>%
  select(name,n_below_45_w_percent,
         gmsc_below_45_w_percent,n_above_55_w_percent,
         gmsc_above_55_w_percent,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  tab_spanner(label="<= 45% Win %",columns=ends_with("45_w_percent")) %>%
  tab_spanner(label=">= 55% Win %",columns=ends_with("55_w_percent")) %>%
  cols_label(
    n_below_45_w_percent="GP",gmsc_below_45_w_percent="GmSc",
    n_above_55_w_percent="GP",gmsc_above_55_w_percent="GmSc"
  )
```

## "The Lights Were Brighter than Expected" Award

biggest PPG decrease from playing against teams below 45% win percentage to playing against teams above 55% win percentage, min 50% of games played and 10 PPG

```{r brighter_than_expected_ppg}
#| tbl-pos: T
#| layout-ncol: 1

opp_win_percent_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2,ppg_total>=10) %>% 
  slice_min(ppg_diff,n=5) %>%
  select(name,n_below_45_w_percent,
         ppg_below_45_w_percent,n_above_55_w_percent,
         ppg_above_55_w_percent,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  tab_spanner(label="<= 45% Win %",columns=ends_with("45_w_percent")) %>%
  tab_spanner(label=">= 55% Win %",columns=ends_with("55_w_percent")) %>%
  cols_label(
    n_below_45_w_percent="GP",ppg_below_45_w_percent="PPG",
    n_above_55_w_percent="GP",ppg_above_55_w_percent="PPG"
  )
```

Alternatively: biggest game score decrease from playing against teams below 45% win percentage to playing against teams above 55% win percentage, min 50% of games played

```{r brighter_than_expected_gmsc}
#| tbl-pos: T
#| layout-ncol: 1

opp_win_percent_ppg_gmsc %>% 
  filter(g_total>=max(g_total)/2) %>% 
  slice_min(gmsc_diff,n=5) %>%
  select(name,n_below_45_w_percent,
         gmsc_below_45_w_percent,n_above_55_w_percent,
         gmsc_above_55_w_percent,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  tab_spanner(label="<= 45% Win %",columns=ends_with("45_w_percent")) %>%
  tab_spanner(label=">= 55% Win %",columns=ends_with("55_w_percent")) %>%
  cols_label(
    n_below_45_w_percent="GP",gmsc_below_45_w_percent="GmSc",
    n_above_55_w_percent="GP",gmsc_above_55_w_percent="GmSc"
  )
```

\newpage

## The "I Gotta Coach AND Ref?" Award (presented by Ace Attorney Phoenix Wright)

Biggest positive difference between actual successful challenges and expected successful challenges (credit to xfinityhomeboy & sleepyfox1312 for the original idea of best success rate on challenges and UBKUBK for the idea of most successful challenges, from <https://official.nba.com/2025-26-nba-coachs-challenge-reviews/>)

```{r challenges_scrape}
nba_bow=bow("https://official.nba.com/",
            user_agent = "Sumitro Datta",force=TRUE,delay=10)
session=nod(nba_bow,path="2025-26-nba-coachs-challenge-reviews/")

coaches_challenge_by_day=scrape(session) %>% 
   html_nodes("a") %>% html_attr("href") %>% tibble() %>%
   rename(links=".") %>%
   distinct(links) %>%
   filter(str_detect(links,"Coachs-Challenge")) %>% slice(1) %>% 
   pull()

# if not working, uncomment below line (download manually)
# coaches_challenge_by_day="Data/2025-26-NBA-Coachs-Challenges-01-12-26.pdf"

challenge_tables=tabulapdf::extract_tables(file=coaches_challenge_by_day)

all_challenges=challenge_tables[[1]] %>% 
  as_tibble(.name_repair = "minimal") %>%
  clean_names()

for (x in 2:length(challenge_tables)){
  new_table=challenge_tables[[x]] %>% 
    as_tibble(.name_repair = "minimal") %>% clean_names()
  all_challenges=bind_rows(all_challenges,new_table)
}
```

```{r challenges_process}
challenges_team_summary=all_challenges %>% 
  mutate(date=mdy(date)) %>%
  #date of first reg season game
  filter(date>=mdy("10/21/2025")) %>%
  group_by(team_challenged) %>% 
  summarize(num_challenges=n(),
            successful_challenges=sum(
              str_detect(challenge_outcome,"Successful"),
              na.rm = TRUE),
            success_rate=successful_challenges/num_challenges)

challenges_team_summary_final=challenges_team_summary %>%
  mutate(
    #remove self from totals
    rest_of_league_success_rate=
      (sum(successful_challenges)-successful_challenges)/
      (sum(num_challenges)-num_challenges),
    expected_successful_challenges=
      rest_of_league_success_rate*num_challenges,
    diff=successful_challenges-expected_successful_challenges
    )

write_csv(challenges_team_summary_final,"Output Data/Challenges Summary.csv")
```

```{r most_success_challenges}
challenges_team_summary_final %>% 
  slice_max(diff,n=5) %>%
  select(team_challenged,num_challenges,successful_challenges,
         success_rate,expected_successful_challenges,diff) %>% gt() %>%
  fmt_percent(columns=success_rate) %>%
  fmt_number(expected_successful_challenges:diff,decimals=2) %>%
  cols_label(team_challenged="Team",num_challenges="Total Challenges",
             successful_challenges="Successful Challenges",
             success_rate="Success Rate",
             expected_successful_challenges="xSuccess",
             diff="Delta") %>%
  as_raw_html()
```

## The "I Did the Math, but Man am I Bad at Math" Award

Biggest negative difference between actual successful challenges and expected successful challenges (credit to livejamie, Necessary_Career_253 and andresalejandro1120 for the original ideas of worst success rate on challenges/most incorrect challenges)

```{r least_success_challenges}
challenges_team_summary_final %>% 
  slice_min(diff,n=5) %>%
  select(team_challenged,num_challenges,successful_challenges,
         success_rate,expected_successful_challenges,diff) %>% gt() %>%
  fmt_percent(columns=success_rate) %>%
  fmt_number(expected_successful_challenges:diff,decimals=2) %>%
  cols_label(team_challenged="Team",num_challenges="Total Challenges",
             successful_challenges="Successful Challenges",
             success_rate="Success Rate",
             expected_successful_challenges="xSuccess",
             diff="Delta") %>% as_raw_html()
```

## The "Just the Tip" Award

best jump ball win percentage, minimum 5 jump balls (credit to livejamie)

```{r ramiro_bentes_pbp}
ramiro_bentes_pbp=readRDS(gzcon(url("https://raw.githubusercontent.com/ramirobentes/nba_pbp_data/refs/heads/main/pbp-final-2026/data.rds")))
```

```{r jump_balls}
#| tbl-pos: T
#| layout-ncol: 1

jump_balls= ramiro_bentes_pbp %>% filter(action_type=="jumpball")

tip_to_team=jump_balls %>% filter(str_detect(description,"Tip to Team")) %>%
  select(game_id,team_home,team_away,team_abb,
         winner=player1_name,description,loser=player2_name)

tip_to_player=jump_balls %>% 
  filter(str_detect(description,"Tip to Team",negate=TRUE)) %>%
  select(game_id,team_home,team_away,team_abb,
         winner=player2_name,description) %>%
  mutate(loser=str_match(description,"vs\\. (.*?):")[,2],
         loser_last_name=word(loser,sep="\\. ",2),
         loser_team_abb=if_else(team_abb==team_home,team_away,team_home))

players_in_game_1=ramiro_bentes_pbp %>% 
  filter(!is.na(team_abb),!is.na(player1_name)) %>%
  distinct(game_id,team_abb,player1_name) %>%
  arrange(game_id,team_abb,player1_name) %>%
  separate_wider_delim(player1_name,delim=" ",
                       names=c("id","first_name","last_name"),
                       too_many="merge",
                       cols_remove = FALSE) %>%
  mutate(first_initial_last_name=
           str_c(str_sub(first_name,end=1),". ",last_name))

tip_to_player_final=left_join(tip_to_player,players_in_game_1,
          by=join_by(loser==first_initial_last_name,
                     loser_team_abb==team_abb,game_id)) %>%
  select(game_id:description,loser=player1_name)

all_jump_balls_final=bind_rows(tip_to_team,tip_to_player_final)

jump_balls_by_player=all_jump_balls_final %>%
  select(winner,loser) %>%
  pivot_longer(winner:loser,values_to = "player") %>% 
  group_by(player) %>%
  summarize(jump_balls_participated=n(),
            wins=sum(name=="winner"),
            losses=sum(name=="loser"),
            win_percent=wins/jump_balls_participated) %>%
  ungroup() %>%
  separate_wider_delim(cols=player,delim=' ',
                       names=c("id","player"),
                       too_many = 'merge')

jump_ball_percentiles=jump_balls_by_player %>% 
  filter(jump_balls_participated>=5) %>%
  mutate(across(c(jump_balls_participated,win_percent),
                percent_rank,
                .names="{.col}_percentile")) %>%
  mutate(avg_percentiles=rowMeans(across(contains("percentile"))))

write_csv(jump_ball_percentiles,"Output Data/Jump Ball Percentiles.csv")

jump_ball_percentiles %>%
  slice_max(avg_percentiles,n=5) %>%
  select(player,jump_balls_participated,win_percent,
         contains("percentile")) %>%
  gt() %>% fmt_percent(columns=-c(player:jump_balls_participated)) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  cols_width(player~px(150),
             win_percent~px(100),
             everything()~px(115))
```

\newpage

## The "Oops, I Dunked It Again" Award (sponsored by Britney Spears, presented by Gary Payton & Shawn Kemp)\*

Most prolific alley-oop duo, both ways (credit to lactardenthusiast for the idea)

```{r alley_oops}
made_alley_oops=ramiro_bentes_pbp %>% 
    filter(str_detect(descriptor,"alley"),shot_result=="Made") %>%
    mutate(p1=pmin(player1_name,player2_name),
         p2=pmax(player1_name,player2_name))

made_alley_oops %>% group_by(team_abb,p1,p2) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Master Baiter" Award (sponsored by Bass Pro Shops & Kleenex)

highest rate of 3-point shooting fouls drawn as percent of total 3-point attempts (PBPStats.com) (credit to OhTheGrandeur for suggesting the change to a rate stat)

```{r master_baiter_scrape}
pbp_stats_data=jsonlite::fromJSON(
  "https://api.pbpstats.com/get-totals/nba?Season=2025-26&SeasonType=Regular+Season&Type=Player&StatType=Totals&Table=FTs")$multi_row_table_data
```

```{r master_baiter_output}
ft_source=pbp_stats_data %>%
  select(name=Name,g=GamesPlayed,team=TeamAbbreviation,
         x3pt_sfd=ThreePtShootingFoulsDrawn,
         x3pt_and_1s=`3pt And 1 Free Throw Trips`,
         FG3M,FG3A) %>%
  replace_na(list(x3pt_sfd=0,x3pt_and_1s=0,FG3M=0,FG3A=0))

rate_based_3pt_fouls_drawn=ft_source %>%
  mutate(x3pt_sfd_tot=x3pt_sfd+x3pt_and_1s,
         rate=x3pt_sfd_tot/FG3A)

rate_based_3pt_fouls_drawn %>% filter(FG3M>=82) %>% 
  slice_max(rate,n=5) %>% 
  select(name,team,x3pt_sfd,x3pt_and_1s,x3pt_sfd_tot,FG3A,rate) %>% 
  gt() %>% fmt_percent(columns=rate) %>% as_raw_html()
```

## The Yeet Award (presented by Jey Uso & [Terry](https://www.youtube.com/shorts/2Bjy5YQ5xPc))

longest distance on 3-pointers attempted (PBPStats)

```{r yeet}
pbp_stats_data %>% filter(FG3M>=82) %>% 
  select(Name,FG3M,FG3A,Avg3ptShotDistance) %>% 
  slice_max(Avg3ptShotDistance,n=5) %>%
  gt() %>% fmt_number(Avg3ptShotDistance,decimals=2) %>% as_raw_html()
```

\newpage

## The "Hug the Line" Award (presented by Cris Carter)

shortest distance on 3-pointers attempted (PBPStats)

```{r hug_the_line}
pbp_stats_data %>% filter(FG3M>=82) %>% 
  select(Name,FG3M,FG3A,Avg3ptShotDistance) %>% 
  slice_min(Avg3ptShotDistance,n=5) %>%
  gt() %>% fmt_number(Avg3ptShotDistance,decimals=2) %>% as_raw_html()
```

## The Wingman Award (presented by Iceman & Maverick)

most assists between teammates, weighing for equality using harmonic mean (PBPStats.com)

```{r wingman_scrape}
assist_combo_data=jsonlite::fromJSON("https://api.pbpstats.com/get-assist-combo-summary/nba?Season=2025-26&SeasonType=Regular%2BSeason")$results
```

```{r wingman_work}
assist_combo_data_both_directions=
  assist_combo_data %>% 
  select(assist_player,assist_player_id,scorer_player,scorer_player_id,team,
         total,total2s,total3s) %>%
  mutate(
    player1 = pmin(assist_player, scorer_player),
    player2 = pmax(assist_player, scorer_player),
    direction = if_else(assist_player == player1, "p1_to_p2", "p2_to_p1")
  ) %>%
  select(player1, player2, team, direction, total, total2s, total3s) %>%
  pivot_wider(
    names_from = direction,
    values_from = c(total, total2s, total3s),
    values_fill = 0
  ) %>%
    mutate(harmonic_mean = 2 * total_p1_to_p2 * total_p2_to_p1 /
                 (total_p1_to_p2 + total_p2_to_p1)
  ) %>%
  arrange(desc(harmonic_mean)) %>%
  mutate(total=total_p1_to_p2+total_p2_to_p1,.before = harmonic_mean) %>%
  mutate(diff_between_sides=abs(total_p1_to_p2-total_p2_to_p1))

write_csv(assist_combo_data_both_directions,"Output Data/Assist Combos.csv")
```

```{r wingman_output}
#| tbl-pos: T
#| layout-ncol: 1

assist_combo_data_both_directions %>%
  select(player1:total_p2_to_p1,total,harmonic_mean) %>%
  slice_max(harmonic_mean,n=5) %>% 
  gt() %>% fmt_number(harmonic_mean,decimals=2) %>%
  tab_spanner(label="Totals",columns = 4:6) %>%
  cols_label(total_p1_to_p2="P1-to-P2",
             total_p2_to_p1="P2-to-P1",
             total="Duo")
```

## The "I'm a Tier 3 Sub!" Award (presented by Pokimane)

biggest discrepancies in assist combinations between teammates (PBPStats.com)

```{r tier_3_sub}
#| tbl-pos: T
#| layout-ncol: 1

assist_combo_data_both_directions %>%
  select(player1:total_p2_to_p1,total,diff_between_sides) %>%
  slice_max(diff_between_sides,n=5) %>% gt() %>%
  tab_spanner(label="Totals",columns = 4:6) %>%
  cols_label(total_p1_to_p2="P1-to-P2",
             total_p2_to_p1="P2-to-P1",
             total="Duo")
```

\newpage

## The Stonks Award

contract overperformance by fewest contract \$ per 1 VORP, excluding players with \<=4 years of experience and less than a minimum salary (credit to memeticengineering for the idea)

```{r stonks_scrape,results='hide'}
contracts=tibble()
for (x in teams) {
  session=nod(bbref_bow,path=paste0("contracts/",x,".html"))
  teams_contracts=scrape(session) %>% 
    html_nodes(xpath='//*[(@id = "contracts")]') %>% 
    html_table() %>% .[[1]] %>% row_to_names(row_number=1) %>% clean_names() %>% 
    filter(age != "",x2025_26 != "") %>% 
    select(player,x2025_26) %>% rename(salary=x2025_26) %>% 
    mutate(salary=parse_number(salary),tm=x)
  contracts=bind_rows(contracts,teams_contracts)
  print(x)
}
```

```{r stonks_df}
salary_performance=left_join(
  contracts,
  #get vorp from advanced
  read_csv("Data/Advanced.csv") %>% 
    filter(season==current_year) %>%
    select(player,player_id,tm=team,vorp),
  by=join_by(player,tm)) %>%
  left_join(
    .,
    #get age & experience from PSI
    read_csv("Data/PLayer Season Info.csv") %>% 
      filter(season==current_year) %>%
      select(player,player_id,age,tm=team,experience),
    by=join_by(player_id,player,tm)) %>%
  select(player,tm,experience,age,salary,vorp) %>% 
  mutate(vorp_per_million=vorp/salary*1000000) %>% 
  mutate(percent_of_cap=salary/154647000) %>% filter(!is.na(vorp))
```

```{r stonks_output}
#remove any player with <=4 years of experience (rookie contract), players w/lower salary than min (10-days)
salary_performance %>% filter(experience > 4,salary>1272870) %>% 
  arrange(desc(vorp_per_million)) %>% 
  select(player,salary,vorp,vorp_per_million) %>% 
  slice_max(vorp_per_million,n=5) %>% gt() %>% 
  fmt_currency(columns=salary,use_subunits = FALSE,
               suffixing = TRUE,decimals=3) %>%
  fmt_number(columns=vorp_per_million,decimals = 4) %>% as_raw_html()
```

Alternatively: salary is at least 5% of the salary cap

```{r stonks_output_min_5_percent_cap}
#remove players whose salary is less than 5% of cap
salary_performance %>% filter(experience > 4,percent_of_cap>0.05) %>% 
  arrange(desc(vorp_per_million)) %>% 
  select(player,salary,vorp,vorp_per_million,percent_of_cap) %>% 
  slice_max(vorp_per_million,n=5) %>% gt() %>% 
  fmt_currency(columns=salary,use_subunits = FALSE,
               suffixing = TRUE,decimals=3) %>%
  fmt_number(columns=vorp_per_million,decimals = 4) %>%
  fmt_percent(columns=percent_of_cap) %>% as_raw_html()

write_csv(salary_performance,"Output Data/Salary Overperformance.csv")
```

## The Rotation Awards

I value depth over one solitary star (it's a rotation after all). My attempt to make this more objective than past years:

-   Get position estimates from Basketball-Reference

    -   players with most minutes at PG & SG: guard

    -   players with most minutes at SG & SF or SF & PF: wing

    -   players with most minutes at PC & C: big

    -   exceptions: all players greater than 7 foot are bigs and all players less than 6 foot are guards

-   Rank teams based on highest positional VORP without positional VORP leader

```{r rotation_heights_scrape,results='hide'}
player_heights=tibble()
for (x in teams) {
  session=nod(bbref_bow,path=paste0("teams/",x,"/",current_year,".html"))
  team_heights=scrape(session) %>% html_nodes(css="#all_roster") %>% 
    html_table() %>% .[[1]] %>% 
    clean_names() %>% select(player:wt)
  player_heights=bind_rows(player_heights,team_heights)
  print(x)
}

final_heights=player_heights %>% 
  mutate(player=str_trim(word(player,sep="\\(TW\\)"))) %>%
  separate(ht,into=c("ht_ft","ht_in"),convert=TRUE) %>% 
  mutate(full_in_ht=12*ht_ft+ht_in)

```

```{r final_rotation_awards}
pos_percents_w_heights=play_by_play %>% 
  select(player,player_id,pos,team:c_percent) %>% 
  filter(str_detect(team,"TM",negate = TRUE)) %>% 
  replace_na(list(pg_percent=0,sg_percent=0,sf_percent=0,
                  pf_percent=0,c_percent=0)) %>%
  mutate(guard_percent=pg_percent+sg_percent,
         wing_small_percent=sg_percent+sf_percent,
         wing_big_percent=sf_percent+pf_percent,
         big_percent=pf_percent+c_percent) %>% 
  pivot_longer(cols=guard_percent:big_percent,names_to="position") %>%
  group_by(player_id,team) %>% slice_max(value,n=1) %>% ungroup() %>% 
  left_join(.,final_heights %>% select(-pos)) %>% 
  mutate(position=case_when(
    full_in_ht>=84~"big_percent",
    full_in_ht<=72~"guard_percent", 
    TRUE~position)) %>%
  #combine wing sizes into one
    mutate(position_2=word(position,sep="_")) %>%
  left_join(
    .,
    read_csv("Data/Advanced.csv") %>% 
      filter(season==current_year, str_detect(team,"TM",negate = TRUE)) %>% 
      select(player,player_id,team,ows:vorp)) %>% 
  #remove players who are both big & small wing
  group_by(player_id,team,position_2) %>% slice(1) %>% ungroup()

#pos_percents_w_heights %>% filter(mp>50) %>% 
#  ggplot(aes(x=full_in_ht,fill=position)) + 
#  geom_bar() + dark_theme_grey()

vorp_leaders_by_team_pos=pos_percents_w_heights %>% 
            group_by(team,position_2) %>% 
  slice_max(vorp,n=1,with_ties=FALSE) %>% 
  mutate(leader=TRUE) %>% ungroup() %>% 
  select(player,player_id,team,position_2,vorp,leader)

pos_percents_w_leaders=left_join(pos_percents_w_heights,
          vorp_leaders_by_team_pos) %>% 
  replace_na(list(leader=FALSE))

write_csv(pos_percents_w_leaders,
          "Output Data/Player Positions.csv")

team_pos_vorps_wo_leader=pos_percents_w_leaders %>% filter(!leader) %>% 
  group_by(team,position_2) %>% 
  summarize(pos_vorp_wo_leader=sum(vorp)) %>% ungroup()

pos_vorps_w_rotations=pos_percents_w_leaders %>%
  filter(mp>=200) %>%
  group_by(team,position_2) %>% 
  arrange(desc(vorp)) %>%
  summarize(rotation=paste0(player,collapse=", ")) %>% 
  ungroup() %>% select(team,position_2,rotation) %>% 
  left_join(.,team_pos_vorps_wo_leader) %>%
  left_join(.,teams_w_conf,by=join_by(team==tm))

write_csv(pos_vorps_w_rotations,"Output Data/Rotations with Positional VORPS.csv")

pos_vorp_leader_filter<-function(position,conf){
  final_df<-pos_vorps_w_rotations %>% 
    filter(position_2==position,conference==conf) %>% 
    slice_max(pos_vorp_wo_leader,n=5) %>% 
    left_join(.,pos_percents_w_leaders %>% 
                select(team,player,position_2,g,mp,vorp)) %>%
    arrange(desc(pos_vorp_wo_leader),desc(vorp)) %>%
    select(team,player,g,mp,vorp,pos_vorp_wo_leader) %>% 
    filter(mp>=200) %>%
    group_by(team,pos_vorp_wo_leader)
}
```

\newpage

### The Best Guard Rotation Award (sponsored by Buckingham Palace)

```{r guards}
#| tbl-pos: T
#| layout-ncol: 2

pos_vorp_leader_filter("guard","East") %>% 
  gt() %>% tab_header("East Guards") %>% tab_options(table.font.size = '70%')

pos_vorp_leader_filter("guard","West") %>% 
  gt() %>% tab_header("West Guards") %>% tab_options(table.font.size = '70%')
```

\newpage

### The Best Wing Rotation Award (presented by Lou Williams, sponsored by Magic City)

```{r wings}
#| tbl-pos: T
#| layout-ncol: 2

pos_vorp_leader_filter("wing","East") %>% 
  gt() %>% tab_header("East Wings") %>% tab_options(table.font.size = '60%')
pos_vorp_leader_filter("wing","West") %>% 
  gt() %>% tab_header("West Wings") %>% tab_options(table.font.size = '60%')
```

\newpage

### The Best Big Rotation Award (presented by Tom Hanks, sponsored by Cadbury and Sex and the City)

```{r bigs}
#| tbl-pos: T
#| layout-ncol: 2

pos_vorp_leader_filter("big","East") %>% 
  gt() %>% tab_header("East Bigs") %>% tab_options(table.font.size = '80%')

pos_vorp_leader_filter("big","West") %>% 
  gt() %>% tab_header("West Bigs") %>% tab_options(table.font.size = '80%')
```
