---
title: "NBA Alt Awards 2026 (Basketball-Reference)"
author: "Sumitro Datta"
format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{typearea}
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=15mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","janitor","gt")
```

```{r data_load}
current_year=2026

advanced=read_csv("Data/Advanced.csv") %>% 
  filter(season==current_year) %>% 
  group_by(player_id) %>% slice_min(team) %>% ungroup()
per_game=read_csv("Data/Player Per Game.csv") %>% 
  filter(season==current_year) %>% 
  group_by(player_id) %>% slice_min(team) %>% ungroup()
totals=read_csv("Data/Player Totals.csv") %>% 
  filter(season==current_year) %>% 
  group_by(player_id) %>% slice_min(team) %>% ungroup()
play_by_play=read_csv("Data/Player Play by Play.csv") %>% 
  filter(season==current_year)

team_summaries=read_csv("Data/Team Summaries.csv") %>% 
  filter(season==current_year) 
team_wl=team_summaries %>% mutate(g_total=w+l) %>% 
  select(season,team=abbreviation,g_total) %>% head(.,-1) %>% 
  add_row(season=current_year,team="2TM",g_total=mean(.$g_total)) %>%
  add_row(season=current_year,team="3TM",g_total=mean(.$g_total))
```

# Awards

## The Real Sixth Man of the Year (presented by Brent Barry)\*

Sixth Man candidates if sixth to ninth on team in minutes played, must have played more than 50% of games and started less than 50% of games, sorted by PPG (credit to KokiriEmerald for the reasoning behind re-implementing the starting criteria)

```{r real_smoy}

smoy_candidates=left_join(advanced,per_game) %>%
  left_join(.,team_wl) %>% 
  mutate(g_percent=g/g_total,gs_percent=gs/g) %>% 
  filter(g_percent>=0.5,
         str_detect(team,"TM",negate=TRUE)) %>% 
  filter(!is.na(mp_per_game)) %>% 
  arrange(desc(mp_per_game)) %>% 
  group_by(team) %>% slice(6:9) %>% ungroup() %>%
  filter(gs_percent<=0.5)

smoy_candidates %>% slice_max(pts_per_game,n=5) %>% 
  select(player,team,g,gs,mp_per_game,pts_per_game) %>% 
  gt() %>% as_raw_html()
```

Alternatively: sorted by VORP

```{r real_smoy_vorp}
smoy_candidates %>% slice_max(vorp,n=5) %>% 
  select(player,team,g,gs,mp_per_game,vorp) %>% 
  gt() %>% as_raw_html()
```

## The Deadshot Award (presented by Ray Allen & Reggie Miller)

best qualifying 3 point percentage, minimum 82 3PM (Basketball-Reference)

```{r deadshot}

# use when players eclipse 82 3PM
totals %>%
  filter(x3p>=82) %>%
  mutate(x3p_per_game=round(x3p/g,1)) %>%
  slice_max(x3p_percent,n=5) %>%
  select(player,team,x3p_per_game,x3p_percent) %>%
  gt() %>% fmt_percent(columns=x3p_percent) %>%
  as_raw_html()

 # per_game %>% left_join(.,team_wl) %>%
 #  filter(x3p_per_game>=1,x3p_per_game*(82-g_total+g)>82) %>%
 #  #filter(x3p_per_game*g>82) %>%
 #  slice_max(x3p_percent,n=5) %>%
 #  select(player,team,x3p_per_game,x3p_percent) %>%
 #  gt() %>% fmt_percent(columns=x3p_percent) %>%
 #  as_raw_html()

```

## The Stormtrooper Award

worst qualifying 2 point percentage, minimum 300 FGM (Basketball-Reference)

```{r stormtrooper}

# use when players eclipse 300 FGM
totals %>%
  filter(fg>=300) %>%
  mutate(x2p_per_game=round(x2p/g,1)) %>%
  slice_min(x2p_percent,n=5) %>%
  select(player,team,x2p_per_game,x2p_percent) %>%
  gt() %>% fmt_percent(columns=x2p_percent) %>%
  as_raw_html()

# per_game %>% left_join(.,team_wl) %>% 
#   filter(fg_per_game*(82-g_total+g)>=300,
#          fg_per_game>=300/82) %>%
#   slice_min(x2p_percent,n=5) %>% 
#   select(player,team,x2p_per_game,x2p_percent) %>% 
#   gt() %>% fmt_percent(columns=x2p_percent) %>%
#   as_raw_html()
```

## The Most Ethical Scorer Award (sponsored by Immanuel Kant)

lowest percentage of points from free throws, minimum 15 PPG and 70% of games played (credit to adc1369 & yeahright17)

```{r ethical_process}
ethical_scoring=totals %>%
  left_join(.,team_wl) %>% mutate(g_percent=g/g_total) %>% 
  mutate(ft_per_game=ft/g,pts_per_game=pts/g,
         ft_percent_of_pts=ft/pts) %>%
  select(player,team,g,g_percent,ft_per_game,pts_per_game,ft_percent_of_pts)

write_csv(ethical_scoring,"Output Data/Ethical Scoring.csv")
```

```{r most_ethical}
ethical_scoring %>%
  filter(g_percent>=58/82,pts_per_game>=15) %>%
  slice_min(ft_percent_of_pts,n=5) %>%
  select(player,team,g,ft_per_game,pts_per_game,ft_percent_of_pts) %>%
  gt() %>%
  fmt_percent(ft_percent_of_pts) %>%
  fmt_number(c(ft_per_game,pts_per_game),decimals = 1) %>%
  as_raw_html()
```

## The Least Ethical Scorer Award (sponsored by ~~Enron~~ ~~Aspiration~~ no one)

highest percentage of points from free throws, minimum 15 PPG and 70% of games played (credit to Banichi-aiji & yeahright17)

```{r least_ethical}
ethical_scoring %>%
  filter(g_percent>=58/82,pts_per_game>=15) %>%
  slice_max(ft_percent_of_pts,n=5) %>%
  select(player,team,g,ft_per_game,pts_per_game,ft_percent_of_pts) %>%
  gt() %>%
  fmt_percent(ft_percent_of_pts) %>%
  fmt_number(c(ft_per_game,pts_per_game),decimals = 1) %>%
  as_raw_html()
```

## The "If He Dies, He Dies" Award (presented by Tom Thibodeau, sponsored by Ivan Drago)

most minutes played per game, minimum 70% of games played (Basketball-Reference) (credit to FurryCrew for the idea)

```{r he_dies_mpg}
per_game %>% 
  left_join(.,team_wl) %>% 
  mutate(g_percent=g/g_total) %>% 
  filter(g_percent>=0.7) %>%
  slice_max(mp_per_game,n=5) %>% 
  select(player,team,g,mp_per_game) %>% 
  gt() %>% as_raw_html()
```

alternatively: most total minutes played (Basketball-Reference) (credit to FrankEMartindale for the idea)

```{r he_dies_total_mp}
totals %>% 
  slice_max(mp,n=5) %>% 
  select(player,team,g,mp) %>% 
  gt() %>% as_raw_html()
```

## The "Black Hole" Award

most FGAs per assist, minimum 50% of games played (credit to Moose4KU for the idea)

```{r black_hole}
field_goals_per_ast=left_join(totals,team_wl) %>% 
  mutate(g_percent=g/g_total) %>% 
  filter(g_percent>=0.5) %>% mutate(fga_per_ast=fga/ast) %>%
  select(player,team,g,mp,fga,ast,fga_per_ast)

write_csv(field_goals_per_ast,"Output Data/FGA Per Assists.csv")

field_goals_per_ast %>%
  slice_max(fga_per_ast,n=5) %>% 
  arrange(desc(fga_per_ast),desc(fga),desc(mp)) %>% 
  gt() %>% 
  fmt_number(fga_per_ast,decimals=4) %>%
  as_raw_html()
```

## The "Hot Potato" Award

fewest FGAs per assist, minimum 50% of games played (credit to Moose4KU for the idea & ajayod for the name)

```{r hot_potato}
field_goals_per_ast %>%
  slice_min(fga_per_ast,n=5) %>% 
  arrange(fga_per_ast,desc(fga),desc(mp)) %>% 
  gt() %>% 
  fmt_number(fga_per_ast,decimals=4) %>%
  as_raw_html()
```

## The "Most Expendable Player" Award (sponsored by the National Basketball Referees Association, presented by Sylvester Stallone)

highest personal fouls per 36 minutes, minimum 50% of games played and 12 MPG (credit to PsychoM & BrightGreenLED for the idea)

```{r most_expendable}
left_join(totals,team_wl) %>% 
  mutate(g_percent=g/g_total) %>% 
  filter(g_percent>=0.5,mp/g>=12) %>% 
  mutate(fouls_per_36_minutes=pf/mp*36) %>% 
  select(player,team,mp,pf,fouls_per_36_minutes) %>%
  slice_max(fouls_per_36_minutes,n=5) %>% 
  gt() %>% 
  fmt_number(fouls_per_36_minutes,decimals=4) %>%
  as_raw_html()
```

## The Weakest Link award (sponsored by Jack Link's Beef Jerky, presented by the 2015 Atlanta Hawks Starting 5)\*

best 5th starter (must have started 50% of a team's games, team must have at least 3 qualifying players) (credit to memeticengineering for the idea)

```{r weakest_link}
best_worst_starter_candidates=
  advanced %>% left_join(.,per_game) %>%
  left_join(.,team_wl) %>% mutate(start_percent=gs/g_total) %>%
  filter(start_percent>=0.5,
         str_detect(team,"TM",negate=TRUE)) %>% 
  mutate(mp_per_game=mp/g,.after="mp") %>% 
  group_by(team) %>%
  filter(n()>2) %>%
  select(player,pos,team,g_total,gs,mp_per_game,vorp) %>%
  slice_min(vorp) %>%
  ungroup()

best_worst_starter_candidates %>% 
  slice_max(vorp,n=5) %>% 
  select(player,pos,team,g_total,gs,mp_per_game,vorp) %>% 
  gt() %>% 
  fmt_number(mp_per_game,decimals=1) %>%
  as_raw_html()
```

## The "[This Game Has Always Been, And Will Always Be, About Buckets](https://www.youtube.com/watch?v=-xYejfYxT4s)" Award

highest points as percentage of counting stats (rebounds, assists, steals, blocks), minimum of 70% of games played

```{r all_about_buckets}
#| tbl-pos: T
#| layout-ncol: 1
pts_as_percent_counting_stats=
  per_game %>% left_join(.,team_wl) %>% 
  mutate(g_percent=g/g_total) %>% filter(g_percent>=0.7) %>% 
  mutate(pts_as_percent_of_other_stats=
           pts_per_game/(pts_per_game+
                           trb_per_game+
                           ast_per_game+
                           stl_per_game+
                           blk_per_game)
         )

pts_as_percent_counting_stats %>% 
  slice_max(pts_as_percent_of_other_stats,n=5) %>% 
  select(player,team,pts_per_game,trb_per_game,ast_per_game,
         stl_per_game,blk_per_game,
         pts_as_percent_of_other_stats) %>% 
  gt() %>% fmt_percent(columns=pts_as_percent_of_other_stats) %>%
  cols_label(pts_per_game="PTS",trb_per_game="TRB",
             ast_per_game="AST",stl_per_game="STL",
             blk_per_game="BLK",
             pts_as_percent_of_other_stats=
               "PTS as % of Other Stats") %>% 
  tab_spanner(label="Per Game",columns=pts_per_game:blk_per_game)

write_csv(pts_as_percent_counting_stats %>% 
            select(player,pts_per_game,trb_per_game,ast_per_game,
                   stl_per_game,blk_per_game,pts_as_percent_of_other_stats),
          "Output Data/Points as Percent of Counting Stats.csv")
```

\newpage

\KOMAoptions{usegeometry,paper=landscape,pagesize}

\recalctypearea

\newgeometry{left=0.5in, right=0.5in, top=0.5in, bottom=0.5in}

## The "Empty Calorie Stats" Award (sponsored by Pop-Tarts)

highest percentile rank within position in usage, descending VORP, descending TS%; minimum of 50% games played (credit to eewap for the idea)

```{r empty_stats}
#| tbl-pos: T
#| layout-ncol: 1
empty_stats_df=advanced %>% 
  left_join(.,team_wl) %>% 
  mutate(g_percent=g/g_total) %>% filter(g_percent>=0.5) %>%
  mutate(pos=word(pos,sep="-")) %>% group_by(pos) %>%
    mutate(ts_rank=percent_rank(desc(ts_percent)),
           usg_rank=percent_rank(usg_percent),
           vorp_rank=percent_rank(desc(vorp)),
           empty_stats=ts_rank+usg_rank+vorp_rank) %>% ungroup()

empty_stats_df %>% 
  slice_max(empty_stats,n=5) %>% 
  select(player,pos,team,ts_percent,usg_percent,vorp,
         ts_rank:empty_stats) %>% 
  gt() %>% 
  fmt_percent(columns=c(ts_percent,ts_rank:empty_stats)) %>%
  cols_label(ts_percent="TS %",usg_percent="USG %",
             ts_rank="TS %",usg_rank="USG %",vorp_rank="VORP") %>%
  tab_spanner(label="Percentile Rank",columns=contains("rank"))

write_csv(empty_stats_df %>% 
            select(player,pos,
                   ts_percent,usg_percent,vorp,
                   ts_rank:empty_stats),
          "Output Data/Empty Stats.csv")
```

## The "Can't Win With These Cats" Award (sponsored by Scar from The Lion King, presented by Kevin Durant in a fake mustache)\*

highest difference in on/off splits in weighted average with and without best, minimum of 50% of games played & 10 MPG (credit to eewap for the idea and ToparBull for the change from median)

```{r high_on_off}
pbp_filtered=play_by_play %>% 
  filter(str_detect(team,"TM",negate=TRUE)) %>% 
  left_join(.,team_wl) %>% mutate(g_percent=g/g_total) %>% 
  filter(g_percent>=0.5,mp/g>=10) %>%
  group_by(team) %>% 
  mutate(tm_avg_weight_npm=sum(
    net_plus_minus_per_100_poss*mp/g)/sum(mp/g)
    ) %>% 
  ungroup()

on_off_leaders=pbp_filtered %>% group_by(team) %>%
  slice_max(net_plus_minus_per_100_poss) %>% ungroup() %>% 
  select(player,team:mp,net_plus_minus_per_100_poss,
         g_percent,tm_avg_weight_npm)

on_off_avg_wo_leader=pbp_filtered %>% group_by(team) %>% 
  arrange(desc(net_plus_minus_per_100_poss)) %>%
  slice(-1) %>% 
  summarize(avg_wo_leader=sum(net_plus_minus_per_100_poss*mp/g)/sum(mp/g))

on_off_diff=left_join(on_off_leaders,on_off_avg_wo_leader) %>% 
  mutate(npm_diff=tm_avg_weight_npm-avg_wo_leader)

on_off_diff %>% slice_max(npm_diff,n=5) %>% 
  select(player:npm_diff) %>% select(-g_percent) %>% 
  gt() %>% 
  cols_label(net_plus_minus_per_100_poss="NPM/100 Poss") %>%
  fmt_number(c(tm_avg_weight_npm,avg_wo_leader,npm_diff),decimals=4) %>%
  as_raw_html()
```

## The "Anchors Aweigh" Award (presented by Ron Burgundy)\*

biggest difference in on/off splits in weighted average with and without worst, minimum of 50% of games played & 10 MPG

```{r low_on_off}
on_off_trailers=pbp_filtered %>% group_by(team) %>%
  slice_min(net_plus_minus_per_100_poss) %>% ungroup() %>% 
  select(player,team:mp,net_plus_minus_per_100_poss,
         g_percent,tm_avg_weight_npm)

on_off_avg_wo_trailer=pbp_filtered %>% group_by(team) %>% 
  arrange(net_plus_minus_per_100_poss) %>%
  slice(-1) %>% 
  summarize(avg_wo_trailer=sum(
    net_plus_minus_per_100_poss*mp)/sum(mp)
    )

on_off_diff_trail=
  left_join(on_off_trailers,on_off_avg_wo_trailer) %>% 
  mutate(npm_diff=tm_avg_weight_npm-avg_wo_trailer)

on_off_diff_trail %>% slice_min(npm_diff,n=5) %>% 
  select(player:npm_diff) %>% select(-g_percent) %>% 
  gt() %>% 
  cols_label(net_plus_minus_per_100_poss="NPM/100 Poss") %>%
  fmt_number(c(tm_avg_weight_npm,avg_wo_trailer,npm_diff),decimals=4) %>%
  as_raw_html()

write_csv(full_join(on_off_diff,on_off_diff_trail) %>% 
            select(player:avg_wo_trailer),"Output Data/On-Off Difference With & Without Extreme Player.csv")
```

\newpage

## The Yoink Award (presented by the Pink Panther)

highest ratio of stocks/36 to defensive fouls/36, minimum 50% of games played (credit to bedlam_au for the idea)

```{r def_stats_and_to_per_36}
def_stats_and_to_per_36=
  left_join(totals,team_wl) %>% mutate(g_percent=g/g_total) %>%
  left_join(.,play_by_play) %>%
  mutate(defensive_fouls=pf-offensive_foul_committed) %>%
  select(player:mp,g_percent,stl,blk,tov,defensive_fouls) %>%
  mutate(stl_per_36=stl/mp*36,
         blk_per_36=blk/mp*36,
         tov_per_36=tov/mp*36,
         stocks_per_36=stl_per_36+blk_per_36,
         def_fouls_per_36=defensive_fouls/mp*36,
         stock_to_def_foul_ratio_per_36=stocks_per_36/def_fouls_per_36,
         stl_plus_tov_per_36=stl_per_36+tov_per_36)

write_csv(def_stats_and_to_per_36,"Output Data/Defensive Stats and Turnovers Per 36.csv")
```

```{r yoink}
def_stats_and_to_per_36 %>%
  filter(g_percent>=0.5) %>%
  slice_max(stock_to_def_foul_ratio_per_36,n=5) %>%
  select(player,team,g,mp,
         stl_per_36,blk_per_36,stocks_per_36,def_fouls_per_36,
         stock_to_def_foul_ratio_per_36) %>%
  gt() %>%
  fmt_number(c(stl_per_36:stock_to_def_foul_ratio_per_36),decimals = 2) %>%
  as_raw_html()
```

## The "I Would've Gotten Away With It If It Wasn't For You Darn Kids" Award (presented by the Scooby Doo Rogues Gallery)

lowest ratio of stocks/36 to defensive fouls/36, minimum 50% of games played (credit to bedlam_au)

```{r would_have_gotten_away_with_it}
def_stats_and_to_per_36 %>%
  filter(g_percent>=0.5) %>%
  slice_min(stock_to_def_foul_ratio_per_36,n=5) %>%
  select(player,team,g,mp,
         stl_per_36,blk_per_36,stocks_per_36,def_fouls_per_36,
         stock_to_def_foul_ratio_per_36) %>%
  gt() %>%
  fmt_number(c(stl_per_36:stock_to_def_foul_ratio_per_36),decimals = 2) %>%
  as_raw_html()
```

## The Lord Giveth & The Lord Taketh Award (presented by Robin Hood)

highest combined sum of steals & turnovers per 36, minimum 50% of games played (credit to PsychoM & Skank_Hunt42 for original idea using steal & turnover percentages)

```{r lord_giveth_and_taketh}
def_stats_and_to_per_36 %>%
  filter(g_percent>=0.5) %>%
  slice_max(stl_plus_tov_per_36,n=5) %>%
  select(player,team,g,mp,stl,tov,stl_per_36,tov_per_36,
         stl_plus_tov_per_36) %>%
  gt() %>%
  fmt_number(c(stl_per_36,tov_per_36,stl_plus_tov_per_36),decimals = 2) %>%
  as_raw_html()
```
