---
title: "NBA Alt Awards Graveyard"
author: "Sumitro Datta"
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","rvest","polite","RKaggle","janitor", "gt")
```

```{r data_load}
current_year=2026

team_summaries=read_csv("Data/Team Summaries.csv") %>% 
  filter(season==current_year) 

bbref_bow=bow("https://www.basketball-reference.com/",
              user_agent = "Sumitro Datta",force=TRUE,delay = 10)
```

```{r full_pbp_load}
current_year=2026
current_year_for_pbp=paste0(current_year-1,"-",str_sub(current_year,-2))
pbp_curr_yr=read_csv(paste0("Data/",current_year_for_pbp,"_pbp.csv")) %>% 
  group_by(gameId) %>% 
  fill(scoreHome,scoreAway) %>%
  mutate(scoreMargin=scoreHome-scoreAway,.after=scoreAway) %>%
  mutate(clutch=(
    (period>4)|
      (period==4 & abs(scoreMargin)<=5 & clock<='PT05')
    )) %>%
  ungroup()
```

# Awards

## The "Don't Check the Box Score" Award

teams that most over performed their Pythagorean record (credit to OhTheGrandeur for the idea)

```{r pyth_overperform}
team_summaries %>% mutate(pyth_diff=w-pw) %>% 
  slice_max(pyth_diff,n=3) %>% select(team,w,pw) %>% gt() %>%
  as_raw_html()
```

## The "Dude, Check the Box Score" Award

teams that most under performed their Pythagorean record (credit to OhTheGrandeur for the idea)

```{r pyth_underperform}
team_summaries %>% mutate(pyth_diff=w-pw) %>% 
  slice_min(pyth_diff,n=3) %>% select(team,w,pw) %>% gt() %>%
  as_raw_html()
```

## The "Fix My Own Mistakes" Award

most shots after rebounding own missed shot (credit to Necessary_Career_253 for the idea)

```{r fix_own_mistake}
pbp_curr_yr_w_before_after=pbp_curr_yr %>% 
  group_by(gameId) %>%
  mutate(next_play=lead(description),
         next_player=lead(personId),
         last_play=lag(description),
         last_player=lag(personId)) %>% ungroup()

shot_after_own_rebound_miss=pbp_curr_yr_w_before_after %>% 
    filter(str_detect(description,"REBOUND")) %>%
  filter(personId==next_player,personId==last_player,
         str_detect(last_play,"MISS")) %>%
  filter(str_detect(next_play,"PTS")) %>% 
  mutate(shot_value=if_else(str_detect(next_play,"3PT"),3,2)) %>% 
  group_by(personId,playerNameI) %>% 
  summarize(fix_with_3=sum(shot_value==3),
            fix_with_2=sum(shot_value==2),
            fixed_shots=n(),
            fixed_points=sum(shot_value)) %>% 
  ungroup()

shot_after_own_rebound_miss %>% slice_max(fixed_points,n=10) %>%
  gt() %>%
  as_raw_html()
```

## The "Human Highlight Reel" Award (presented by Dominique Wilkins)\*

players that had the most games with 2+ dunks (credit to frosiano for the idea)

```{r two_plus_dunks}
made_dunks=pbp_curr_yr %>% 
    filter(str_detect(description,"Dunk"),shotResult=="Made")

made_dunks %>% group_by(gameId,personId,playerNameI) %>% 
  summarize(dunks_in_game=n()) %>%
  filter(dunks_in_game>=2) %>% ungroup() %>% 
  group_by(personId,playerNameI) %>% 
  summarize(games_w_2plus_dunks=n()) %>% 
  ungroup() %>%
  slice_max(games_w_2plus_dunks,n=10) %>% gt() %>%
  as_raw_html()
```

## The "Most Spectacular Self-Destruction" Award

furthest deviation from preseason record over/unders (credit to Hockeyguy928 for the name)

```{r self_destruct}
session=nod(bbref_bow,path=paste0("leagues/NBA_",current_year,
                                  "_preseason_odds.html"))
preseason_odds=scrape(session) %>% 
  html_nodes(xpath='//*[(@id = "NBA_preseason_odds")]') %>% 
  html_table() %>% .[[1]]

preseason_odds %>% clean_names() %>% filter(str_detect(result,"under")) %>%
  mutate(wins=as.numeric(word(result,sep="-")),
         losses=as.numeric(word(word(result,sep="-",2),sep=" ")),
         pace=round((wins/(wins+losses))*82,0),
         diff=w_l_o_u-pace) %>% 
  slice_max(diff,n=3) %>% select(-x) %>% gt()
```

## The "Boneheaded Foul" Award (presented by Sideshow Mel) \[TO-DO\]

```{r boneheaded,eval=FALSE}
View(
  pbp_curr_yr %>% 
    #one possession games
    filter(abs(scoreMargin)<=3) %>%
    #check to see who is leading
    mutate(leading=if_else(scoreMargin<0,"v","h")) %>%
    #last 30 seconds of 4th quarter and any overtime
    filter(period>3 & clock<='PT00M30') %>%
    filter(actionType=="Foul",leading==location)
)
```

## The "I'm Gonna Get Mine" Award

biggest PPG increase from losses to wins, min 50% of games played, 10 PPG and more losses than wins (credit to Drummallumin for the idea)

```{r box_scores}
x2025_2026_box_scores=
  get_dataset("eoinamoore/historical-nba-data-and-player-box-scores") %>%
  .[[5]] %>% filter(gameDateTimeEst>=mdy("10/21/2025")) %>% 
  filter(!is.na(numMinutes) & !is.na(win)) %>% 
  mutate(game_score=
           points+fieldGoalsMade*0.4-0.7*fieldGoalsAttempted-
           0.4*(freeThrowsAttempted-freeThrowsMade)+
           0.7*reboundsOffensive+0.3*reboundsDefensive+steals+
           0.7*assists+0.7*blocks-0.4*foulsPersonal-turnovers)
```

```{r win_loss_split_process}
win_loss_ppg_gmsc=x2025_2026_box_scores %>%
  group_by(personId,firstName,lastName,win) %>% 
  summarize(n=n(),points=sum(points),game_score=sum(game_score)) %>%
  mutate(win=if_else(win==0,"loss","win"),
         name=paste(firstName,lastName)) %>%
  pivot_wider(names_from=win,values_from = c(n,points,game_score),
              values_fill = 0) %>%
  mutate(g_total=n_loss+n_win,
         pts_total=points_loss+points_win,
         ppg_loss=points_loss/n_loss,
         ppg_win=points_win/n_win,
         ppg_total=pts_total/g_total,
         ppg_diff=ppg_loss-ppg_win,
         gmsc_loss=game_score_loss/n_loss,
         gmsc_win=game_score_win/n_win,
         gmsc_diff=gmsc_loss-gmsc_win) %>%
  ungroup()
```

```{r gonna_get_mine_ppg}
win_loss_ppg_gmsc %>% filter(
  g_total>=(win_loss_ppg_gmsc %>% 
    slice_max(g_total,n=1,with_ties = FALSE) %>% pull(g_total))/2,
  n_loss>=n_win,
  ppg_total>=10
) %>% slice_max(ppg_diff,n=5) %>%
  select(name,n_loss,ppg_loss,n_win,ppg_win,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  as_raw_html()
```

Alternatively: biggest game score increase from road to home, min 50% of games played and more losses than wins

```{r gonna_get_mine_gmsc}
win_loss_ppg_gmsc %>% filter(
  g_total>=(win_loss_ppg_gmsc %>% 
    slice_max(g_total,n=1,with_ties = FALSE) %>% pull(g_total))/2,
  n_loss>=n_win
) %>% slice_max(gmsc_diff,n=5) %>%
  select(name,n_loss,gmsc_loss,n_win,gmsc_win,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  as_raw_html()
```

## The "I Put the Team on My Back Doe" (presented by [Greg Jennings](https://www.youtube.com/watch?v=1P0yfq2wDvU))

biggest PPG decrease from wins to losses, min 50% of games played, 10 PPG and more wins than losses (credit to Drummallumin for the idea)

```{r team_on_my_back_ppg}
win_loss_ppg_gmsc %>% filter(
  g_total>=(win_loss_ppg_gmsc %>% 
    slice_max(g_total,n=1,with_ties = FALSE) %>% pull(g_total))/2,
  n_loss<=n_win,
  ppg_total>=10
) %>% slice_min(ppg_diff,n=5) %>%
  select(name,n_win,ppg_win,n_loss,ppg_loss,ppg_diff) %>%
  gt() %>% 
  fmt_number(contains("ppg"),decimals=2) %>%
  as_raw_html()
```

Alternatively: biggest game score decrease from wins to losses, min 50% of games played and more wins than losses

```{r team_on_my_back_gmsc}
win_loss_ppg_gmsc %>% filter(
  g_total>=(win_loss_ppg_gmsc %>% 
    slice_max(g_total,n=1,with_ties = FALSE) %>% pull(g_total))/2,
  n_loss<=n_win
) %>% slice_min(gmsc_diff,n=5) %>%
  select(name,n_win,gmsc_win,n_loss,gmsc_loss,gmsc_diff) %>%
  gt() %>% 
  fmt_number(contains("gmsc"),decimals=2) %>%
  as_raw_html()
```
