---
title: "NBA Alt Awards 2026 (Scraping)"
author: "Sumitro Datta"
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","rvest","polite","janitor", "gt")
```

```{r data_load}
current_year=2026
```

# Awards

## The "Master Baiter" Award (sponsored by Bass Pro Shops & Kleenex)

highest rate of 3-point shooting fouls drawn as percent of total 3-point attempts (PBPStats.com) (credit to OhTheGrandeur for suggesting the change to a rate stat)

```{r master_baiter_scrape}
pbp_stats_data=jsonlite::fromJSON(
  "https://api.pbpstats.com/get-totals/wnba?Season=2026&SeasonType=Regular+Season&Type=Player&StatType=Totals&Table=FTs")$multi_row_table_data
```

```{r master_baiter_output}
ft_source=pbp_stats_data %>%
  select(name=Name,g=GamesPlayed,team=TeamAbbreviation,
         x3pt_sfd=ThreePtShootingFoulsDrawn,
         x3pt_and_1s=`3pt And 1 Free Throw Trips`,
         FG3M,FG3A) %>%
  replace_na(list(x3pt_sfd=0,x3pt_and_1s=0,FG3M=0,FG3A=0))

rate_based_3pt_fouls_drawn=ft_source %>%
  mutate(x3pt_sfd_tot=x3pt_sfd+x3pt_and_1s,
         rate=x3pt_sfd_tot/FG3A)

rate_based_3pt_fouls_drawn %>% filter(FG3M>=30) %>% 
  slice_max(rate,n=5) %>% 
  select(name,team,x3pt_sfd,x3pt_and_1s,x3pt_sfd_tot,FG3A,rate) %>% 
  gt() %>% fmt_percent(columns=rate) %>% as_raw_html()
```

## The Yeet Award (presented by Jey Uso & [Terry](https://www.youtube.com/shorts/2Bjy5YQ5xPc))

longest distance on 3-pointers attempted, min 30 3PM (PBPStats)

```{r yeet}
pbp_stats_data %>% filter(FG3M>=30) %>% 
  select(Name,FG3M,FG3A,Avg3ptShotDistance) %>% 
  slice_max(Avg3ptShotDistance,n=5) %>%
  gt() %>% fmt_number(Avg3ptShotDistance,decimals=2) %>% as_raw_html()
```

## The "Hug the Line" Award

shortest distance on 3-pointers attempted, min 30 3PM (PBPStats.com)

```{r hug_the_line}
pbp_stats_data %>% filter(FG3M>=30) %>% 
  select(Name,FG3M,FG3A,Avg3ptShotDistance) %>% 
  slice_min(Avg3ptShotDistance,n=5) %>%
  gt() %>% fmt_number(Avg3ptShotDistance,decimals=2) %>% as_raw_html()
```

## The Wingwoman Award (presented by Thelma & Louise)

most assists between teammates, weighing for equality using harmonic mean (PBPStats.com)

```{r wingwoman_scrape}
assist_combo_data=jsonlite::fromJSON("https://api.pbpstats.com/get-assist-combo-summary/wnba?Season=2026&SeasonType=Regular%2BSeason")$results
```

```{r wingwoman_work}
assist_combo_data_both_directions=
  assist_combo_data %>% 
  select(assist_player,assist_player_id,scorer_player,scorer_player_id,team,
         total,total2s,total3s) %>%
  mutate(
    player1 = pmin(assist_player, scorer_player),
    player2 = pmax(assist_player, scorer_player),
    direction = if_else(assist_player == player1, "p1_to_p2", "p2_to_p1")
  ) %>%
  select(player1, player2, team, direction, total, total2s, total3s) %>%
  pivot_wider(
    names_from = direction,
    values_from = c(total, total2s, total3s),
    values_fill = 0
  ) %>%
    mutate(harmonic_mean = 2 * total_p1_to_p2 * total_p2_to_p1 /
                 (total_p1_to_p2 + total_p2_to_p1)
  ) %>%
  arrange(desc(harmonic_mean)) %>%
  mutate(total=total_p1_to_p2+total_p2_to_p1,.before = harmonic_mean) %>%
  mutate(diff_between_sides=abs(total_p1_to_p2-total_p2_to_p1))
```

```{r wingwoman_output}
#| tbl-pos: T
#| layout-ncol: 1

assist_combo_data_both_directions %>%
  select(player1:total_p1_to_p2,total,harmonic_mean) %>%
  slice_max(harmonic_mean,n=5) %>% 
  gt() %>% fmt_number(harmonic_mean,decimals=2) %>%
  tab_spanner(label="Totals",columns = 4:6) %>%
  cols_label(total_p1_to_p2="P1-to-P2",
             total_p2_to_p1="P2-to-P1",
             total="Duo")
```

## The "I'm a Tier 3 Sub!" Award (presented by Pokimane)

biggest discrepancies in assist combinations between teammates (PBPStats.com)

```{r tier_3_sub}
#| tbl-pos: T
#| layout-ncol: 1

assist_combo_data_both_directions %>%
  select(player1:total_p1_to_p2,total,diff_between_sides) %>%
  slice_max(diff_between_sides,n=5) %>% gt() %>%
  tab_spanner(label="Totals",columns = 4:6) %>%
  cols_label(total_p1_to_p2="P1-to-P2",
             total_p2_to_p1="P2-to-P1",
             total="Duo")
```
