---
title: "NBA Alt Awards 2026 (hoopR)"
author: "Sumitro Datta"
format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{typearea}
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","hoopR","janitor","gt")
```

# Awards

## The Spark Plug Award (sponsored by Lt. Surge, presented by American Express CEO Stephen J Squeri)

Most charges drawn per 36 minutes, minimum 70% of games played (credit to morron88 for the idea to separate charges & loose balls in 2020)

```{r hustle_full}
compact_standings=nba_leaguestandingsv3()$Standings %>% 
  type_convert() %>%
  select(TeamID,TeamCity,TeamName,Conference,Division,WINS,LOSSES) %>% 
  mutate(TeamGP=WINS+LOSSES,TeamFullName=paste(TeamCity,TeamName))

player_bio=
  nba_leaguedashplayerbiostats()$LeagueDashPlayerBioStats %>% 
  select(PLAYER_ID:GP) %>%
  type_convert()

player_w_gp_percentages=left_join(
  player_bio,
  compact_standings %>% select(contains("Team")),
  by=join_by(TEAM_ID==TeamID)
  ) %>%
  mutate(G_PERCENT=GP/TeamGP)

rm(player_bio,compact_standings)

hustle=nba_leaguehustlestatsplayer(per_mode="Totals")$HustleStatsPlayer %>% 
  type_convert() %>% select(-contains("PCT"))
hustle_w_gp_qualify=left_join(hustle,
                              player_w_gp_percentages %>% 
                                select(PLAYER_ID,G_PERCENT)) %>%
  relocate(G_PERCENT,.after=G) %>%
  mutate(
    across(-c(PLAYER_ID:MIN),.fns=~./MIN*36,.names="{.col}_per_36")
    )
```

```{r charges_drawn}
hustle_w_gp_qualify %>% filter(G_PERCENT>=0.7) %>%
  slice_max(CHARGES_DRAWN_per_36,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,MIN,
         CHARGES_DRAWN,CHARGES_DRAWN_per_36) %>% 
  gt() %>%
  fmt_number(CHARGES_DRAWN_per_36,decimals=4) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  as_raw_html()
```

## The Most Loose Balls Recovered Award (sponsored by Hungry Hungry Hippos, presented by Dennis Rodman & [Nene's doctor](https://www.espn.com/nba/news/story?id=3197423))

Per 36 minutes, minimum 70% of games played

```{r loose_balls}
hustle_w_gp_qualify %>% filter(G_PERCENT>=0.7) %>%
  slice_max(LOOSE_BALLS_RECOVERED_per_36,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,MIN,
         LOOSE_BALLS_RECOVERED,LOOSE_BALLS_RECOVERED_per_36) %>% 
  gt() %>% fmt_number(LOOSE_BALLS_RECOVERED_per_36,decimals=4) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  as_raw_html()
```

\newpage

## The Plexiglass Award

most deflections per 36 minutes, minimum 70% of games played

```{r deflections}
hustle_w_gp_qualify %>% filter(G_PERCENT>=0.7) %>%
  slice_max(DEFLECTIONS_per_36,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,MIN,
         DEFLECTIONS,DEFLECTIONS_per_36) %>%
  gt() %>% fmt_number(DEFLECTIONS_per_36,decimals=4) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  as_raw_html()
```

## The Wes Unseld Memorial Brick Wall Award

most points generated by screen assists per 36 minutes, minimum 70% of games played

```{r screen_ast_pts}
hustle_w_gp_qualify %>% filter(G_PERCENT>=0.7) %>%
  slice_max(SCREEN_AST_PTS_per_36,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,MIN,
         SCREEN_AST_PTS,SCREEN_AST_PTS_per_36) %>% 
  gt() %>% fmt_number(SCREEN_AST_PTS_per_36,decimals=4) %>%
  cols_label_with(fn=~str_replace_all(.,"_"," ")) %>%
  as_raw_html()
```

\newpage

\KOMAoptions{usegeometry,paper=landscape,pagesize}

\recalctypearea

\newgeometry{left=0.5in, right=0.5in, top=0.5in, bottom=0.5in}

## The "He Trick Y'All, Running Around, Doing Nothing" Award (sponsored by Russell Westbrook, presented by Tony Snell)

Lowest sum of per-36 percentile ranks in the following: charges, contested shots, deflections, defensive boxouts, defensive loose balls recovered (minimum 50% of games played)

```{r hustle_ranks_bottom}
#| tbl-pos: T
#| layout-ncol: 1

hustle_50_percent_gp=hustle_w_gp_qualify %>% filter(G_PERCENT>=0.5) %>% 
  select(PLAYER_NAME,MIN,contains("CHARGES"),
         contains("2PT"),contains("3PT"),
         contains("DEFLECTIONS"),
         contains("DEF_BOX"),contains("DEF_LOOSE")) %>%
  relocate(contains("per_36"),.after=everything()) %>%
  mutate(across(contains("per_36"),percent_rank,
                .names = "{col}_percentile_rk")) %>%
  mutate(sum=rowSums(across(contains("percentile_rk"))))

write_csv(hustle_50_percent_gp,"Output Data/Hustle Ranks.csv")

hustle_50_percent_gp %>% slice_min(sum,n=5) %>% 
  select(player=PLAYER_NAME,MIN,contains("percentile_rk"),sum) %>%
  gt() %>% 
  fmt_percent(columns=-c(player,MIN)) %>%
  tab_spanner(label="Per 36 Percentile Rank",columns=3:8) %>%
  cols_label_with(
    fn=~str_replace_all(
      str_replace_all(.,"per_36_percentile_rk",""),"_"," "
      )
    ) %>%
  cols_width(-c(player,MIN,sum)~px(115))
```

## The "Got that Dawg in Him" Award (presented by Air Bud)\*

Highest sum of per-36 percentile ranks in the following: charges, contested shots, deflections, defensive boxouts, defensive loose balls recovered (minimum 50% of games played) (credit to memeticengineering for the idea)

```{r hustle_ranks_top}
#| tbl-pos: T
#| layout-ncol: 1

hustle_50_percent_gp %>% slice_max(sum,n=5) %>% 
  select(player=PLAYER_NAME,MIN,contains("percentile_rk"),sum) %>%
  gt() %>% 
  fmt_percent(columns=-c(player,MIN)) %>%
  tab_spanner(label="Per 36 Percentile Rank",columns=3:8) %>%
  cols_label_with(
    fn=~str_replace_all(
      str_replace_all(.,"per_36_percentile_rk",""),"_"," "
      )
    ) %>%
  cols_width(-c(player,MIN,sum)~px(115))
```

\newpage

\KOMAoptions{paper=portrait,pagesize}

\recalctypearea

\restoregeometry

## The BBQ Chicken Award (presented by Shaquille O'Neal)

worst isolation defenders (high frequency & high points per possession allowed) (credit to TitanTigers & RTLT512)

```{r bbq_chicken}
nba_synergyplaytypes(play_type = "Isolation",
                     per_mode = "Totals",
                     type_grouping = "Defensive")$SynergyPlayType %>% 
  type_convert() %>% 
  mutate(freq_percentile=percent_rank(POSS_PCT),
         avg_rank=(freq_percentile+(1-PERCENTILE))/2) %>%
  slice_max(avg_rank,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         POSS,POSS_PCT,PPP,freq_percentile,PPP_PERCENTILE=PERCENTILE,
         avg_rank) %>%
  gt() %>%
  fmt_percent(c(POSS_PCT,contains("percentile"),avg_rank)) %>%
  as_raw_html()
```

## The Rotisserie Chicken Award (sponsored by Costco)

worst PNR ball handler defenders (high frequency & high points per possession allowed) (credit to RTLT512 & altu_0002)

```{r rotisserie_chicken}

nba_synergyplaytypes(play_type = "PRBallHandler",
                     per_mode = "Totals",
                     type_grouping = "Defensive")$SynergyPlayType %>% 
  type_convert() %>% 
  mutate(freq_percentile=percent_rank(POSS_PCT),
         avg_rank=(freq_percentile+(1-PERCENTILE))/2) %>%
  slice_max(avg_rank,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         POSS,POSS_PCT,PPP,freq_percentile,PPP_PERCENTILE=PERCENTILE,
         avg_rank) %>% 
  gt() %>%
  fmt_percent(c(POSS_PCT,contains("percentile"),avg_rank)) %>%
  as_raw_html()
```

## The "Would You Mind Coming with Us, Sir?" Award (sponsored by the TSA, presented by Paul Blart)

worst screen defenders (high frequency & high points per possession allowed) (credit to blackjack_trial)

```{r tsa_screen}
nba_synergyplaytypes(play_type = "OffScreen",
                     per_mode = "Totals",
                     type_grouping = "Defensive")$SynergyPlayType %>% 
  type_convert() %>% 
  mutate(freq_percentile=percent_rank(POSS_PCT),
         avg_rank=(freq_percentile+(1-PERCENTILE))/2) %>%
  slice_max(avg_rank,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         POSS,POSS_PCT,PPP,freq_percentile,PPP_PERCENTILE=PERCENTILE,
         avg_rank) %>%
  gt() %>%
  fmt_percent(c(POSS_PCT,contains("percentile"),avg_rank)) %>%
  as_raw_html()
```

\newpage

\KOMAoptions{usegeometry,paper=landscape,pagesize}

\recalctypearea

\newgeometry{left=0.5in, right=0.5in, top=0.5in, bottom=0.5in}

## The Trickshot Grenadier Award (presented by Dude Perfect)

Highest sum of percentile ranks in FGA, FGA frequency & eFG% on shots with 4 seconds or less on the shotclock (credit to BehavioralSink & Bylanta for the idea)

```{r trickshot_grenadier}
#| tbl-pos: T
#| layout-ncol: 1

trickshot_grenades=
  nba_leaguedashplayerptshot(
    shot_clock_range="4-0 Very Late")$LeagueDashPTShots %>% 
  type_convert() %>% 
  select(PLAYER_ID:EFG_PCT) %>% 
  mutate(across(c(FGA_FREQUENCY,FGA,EFG_PCT),
              percent_rank,
              .names="{.col}_percentile")) %>%
  mutate(sum_percentiles=rowSums(across(contains("percentile"))))

write_csv(trickshot_grenades,"Output Data/Trickshot Grenadier.csv")

trickshot_grenades %>%
  slice_max(sum_percentiles,n=5) %>%
  select(PLAYER_NAME,FGA_FREQUENCY,FGA,EFG_PCT,contains("percentile")) %>%
  gt() %>%
  fmt_percent(columns=-c(PLAYER_NAME,FGA)) %>%
  cols_label(FGA_FREQUENCY_percentile="FGA_FREQ",
                  FGA_percentile="FGA",
                  EFG_PCT_percentile="EFG_PCT") %>%
  tab_spanner(label="Percentile Rank",
              columns=FGA_FREQUENCY_percentile:EFG_PCT_percentile)
```

## The "Paint Allergy" Award

players w/highest % of shots from outside paint, min 150 FGA (credit to frosiano for the original idea of highest percentage of 3FGA of total FGA, and to Drummallumin for the revised idea of all shots outside of 15 feet)

```{r paint_allergy}
non_paint_fga=nba_leaguedashplayershotlocations()$ShotLocations %>%
  type_convert() %>%
  select(-c(Corner_3_FGA:Corner_3_FG_PCT)) %>%
  replace(is.na(.),0) %>%
  mutate(tot_fga=rowSums(select(.,contains("FGA"))),
         tot_fgm=rowSums(select(.,contains("FGM"))),
         non_paint_fga=
           tot_fga-Restricted_Area_FGA-In_The_Paint_Non_RA_FGA,
         non_paint_fga_percentage=non_paint_fga/tot_fga)

left_join(non_paint_fga,player_w_gp_percentages) %>%
  filter(tot_fgm/GP>=150/82,G_PERCENT>=0.5) %>%
  slice_max(non_paint_fga_percentage,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,tot_fga,
         non_paint_fga,non_paint_fga_percentage) %>% gt() %>% 
  fmt_percent(non_paint_fga_percentage) %>%
  as_raw_html()
```

## The "Lumberjack" Award (sponsored by Paul Bunyan)

players w/highest % of shots from inside paint, max height of 6'3" & min 150 FGA (credit to Drummallumin for the idea)

```{r lumberjack}
left_join(non_paint_fga,player_w_gp_percentages) %>%
  filter(tot_fgm/GP>=150/82,G_PERCENT>=0.5,
         PLAYER_HEIGHT_INCHES<=6*12+3) %>%
  mutate(paint_fga=tot_fga-non_paint_fga,
         paint_fga_percentage=1-non_paint_fga_percentage) %>%
  slice_max(paint_fga_percentage,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,PLAYER_HEIGHT,
         tot_fga,paint_fga,paint_fga_percentage) %>% gt() %>% 
  fmt_percent(paint_fga_percentage) %>%
  as_raw_html()
```

\newpage

\KOMAoptions{paper=portrait,pagesize}

\recalctypearea

\restoregeometry

## The "FUCK OUTTA HERE, I GOT THAT SHIT" Award (presented by Carmelo Anthony)

Lowest contested rebound percentage, minimum 50% of games played

```{r rebounding}
reb_w_gp_qualify=left_join(
  nba_playerdashptreb(
  team_id=0,player_id=0)$OverallRebounding %>%
  type_convert(),
  player_w_gp_percentages %>% 
    select(PLAYER_NAME,TEAM_ABBREVIATION,PLAYER_ID,G_PERCENT,
           PLAYER_HEIGHT,PLAYER_HEIGHT_INCHES)
  )
```

```{r melo_rebound}
reb_w_gp_qualify %>%
  filter(G_PERCENT>=0.5) %>% slice_min(C_REB_PCT,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         PLAYER_HEIGHT,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

alternatively: restricting to players \> 6 foot 6 inches in height

```{r rebound_above_66}
reb_w_gp_qualify %>% filter(PLAYER_HEIGHT_INCHES>6*12+6) %>% 
  filter(G_PERCENT>=0.5) %>% slice_min(C_REB_PCT,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         PLAYER_HEIGHT,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

## The "Glass Cleaner" Award (presented by Dennis Rodman, sponsored by Windex)

Highest contested rebound percentage, minimum 50% of games played

```{r glass_cleaner}
reb_w_gp_qualify %>% 
  filter(G_PERCENT>=0.5) %>% slice_max(C_REB_PCT,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         PLAYER_HEIGHT,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

alternatively: restricting to players \< 6 foot 7 inches in height

```{r rebound_below_67}
reb_w_gp_qualify %>% filter(PLAYER_HEIGHT_INCHES<6*12+7) %>% 
  filter(G_PERCENT>=0.5) %>% slice_max(C_REB_PCT,n=5) %>% 
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         PLAYER_HEIGHT,C_REB_PCT) %>% gt() %>%
  fmt_percent(columns=C_REB_PCT) %>%
  as_raw_html()
```

## The "Fine, I'll Do It Myself" Award (sponsored by Thanos, presented by Allen Iverson)

Highest percentage of unassisted field goals, minimum 50% of games played

```{r field_goals_ast_and_uast}
scoring_w_gp_qualify=left_join(
  nba_leaguedashplayerstats(
    measure_type = "Scoring")$LeagueDashPlayerStats %>% type_convert(),
  player_w_gp_percentages) %>%
  filter(G_PERCENT>=0.5)
```

```{r do_it_myself}
scoring_w_gp_qualify %>%
  slice_max(PCT_UAST_FGM,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         GP,MIN,FGM,PCT_UAST_FGM) %>% gt() %>%
  fmt_percent(PCT_UAST_FGM) %>%
  as_raw_html()
```

## The "You Gotta Feed Me" Award (presented by Joey Chestnut & Marcin Gortat)

Highest percentage of assisted field goals, minimum 50% of games played and 1 FGM per game

```{r gotta_feed}
scoring_w_gp_qualify %>%
  filter(FGM/GP>=1) %>%
  slice_max(PCT_AST_FGM,n=5) %>%
  select(player=PLAYER_NAME,team=TEAM_ABBREVIATION,
         GP,MIN,FGM,PCT_AST_FGM) %>% gt() %>%
  fmt_percent(PCT_AST_FGM) %>%
  as_raw_html()
```
