---
title: "NBA Alt Awards 2026 (NBA PBP)"
author: "Sumitro Datta"
format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{typearea}
editor: visual
execute:
  echo: false
  warning: false
  message: false
geometry:
  - top=20mm
  - left=15mm
  - right=15mm
---

```{r load_pkgs, results="hide"}
pacman::p_load("tidyverse","hoopR","janitor","gt")
```

```{r full_pbp_load}
current_year=2026
current_year_for_pbp=paste0(current_year-1,"-",str_sub(current_year,-2))
pbp_curr_yr=read_csv(paste0("Data/",current_year_for_pbp,"_pbp.csv")) %>% 
  group_by(gameId) %>% 
  fill(scoreHome,scoreAway) %>%
  mutate(scoreMargin=scoreHome-scoreAway,.after=scoreAway) %>%
  mutate(clutch=(
    (period>4)|
      (period==4 & abs(scoreMargin)<=5 & clock<='PT05')
    )) %>%
  ungroup() %>%
  # remove allstar weekend
  filter(str_starts(as.character(gameId),'3',negate = TRUE))
```

```{r hoopr_info}
compact_standings=nba_leaguestandingsv3()$Standings %>% 
  type_convert() %>%
  select(TeamID,TeamCity,TeamName,Conference,Division,WINS,LOSSES) %>% 
  mutate(TeamGP=WINS+LOSSES,TeamFullName=paste(TeamCity,TeamName))

player_bio=
  nba_leaguedashplayerbiostats()$LeagueDashPlayerBioStats %>% 
  select(PLAYER_ID:GP) %>%
  type_convert()

team_info=left_join(compact_standings,
          player_bio %>% 
            group_by(TEAM_ID) %>% 
            slice(1) %>% 
            select(TEAM_ID,TEAM_ABBREVIATION), 
          by=join_by(TeamID==TEAM_ID))

player_stats=nba_leaguedashplayerstats()$LeagueDashPlayerStats %>% 
  type_convert()

player_w_gp_percentages=left_join(
  player_bio,
  compact_standings %>% select(contains("Team")),
  by=join_by(TEAM_ID==TeamID)
  ) %>%
  left_join(.,
            player_stats %>% select(PLAYER_ID,MIN)) %>%
  mutate(G_PERCENT=GP/TeamGP)

rm(compact_standings,player_bio,player_stats)
```

# Awards

## The "Imma Make 'Em Both" Award (presented by Diar DeRozan & Grant Williams)\*

biggest decline from non-clutch FT% to clutch FT%, min 10 clutch FTA (credit to Necessary_Career_253 for the idea and midnightgreen29 for the name)

```{r clutch_ft_worst}
free_throws=pbp_curr_yr %>% 
  filter(str_detect(description,"Free Throw"))

ft_div_by_clutch=free_throws %>% 
  group_by(personId,playerNameI,clutch) %>%
  summarize(fts_attempted=n(),
            fts_missed=sum(str_detect(description,"MISS"),
                                  na.rm=TRUE),
            ft_percent=1-fts_missed/fts_attempted) %>% 
  ungroup() %>%
  mutate(clutch=if_else(clutch,"clutch","non_clutch")) %>%
  pivot_wider(names_from=clutch,
              values_from=fts_attempted:ft_percent,
              values_fill=0,
              names_glue = "{clutch}_{.value}") %>%
  mutate(differential=clutch_ft_percent-non_clutch_ft_percent) %>%
  left_join(.,player_w_gp_percentages %>% 
              select(personId=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId) %>% select(-playerNameI)
  
write_csv(ft_div_by_clutch,
          "Output Data/Clutch Free Throw Differential.csv")

ft_div_by_clutch %>% filter(clutch_fts_attempted>=10) %>%
  slice_min(differential,n=5) %>% 
  select(player_name,clutch_fts_attempted,
         clutch_ft_percent,non_clutch_ft_percent,differential) %>%
  gt() %>% 
  fmt_percent(columns=-c(player_name,clutch_fts_attempted)) %>%
  as_raw_html()
```

## The "Ice, Ice, Baby" Award (sponsored by Hisense, presented by Vanilla Ice)\*

biggest improvement from non-clutch FT% to clutch FT%, min 10 clutch FTA

```{r clutch_ft_best}
ft_div_by_clutch %>% filter(clutch_fts_attempted>=10) %>%
  slice_max(differential,n=5) %>% 
  select(player_name,clutch_fts_attempted,
         clutch_ft_percent,non_clutch_ft_percent,differential) %>%
  gt() %>% 
  fmt_percent(columns=-c(player_name,clutch_fts_attempted)) %>%
  as_raw_html()
```

## The "Chicken Supplier" Award (sponsored by Los Pollos Hermanos)

most pairs of free throws missed by a visiting player in the second half (credit to livejamie for the idea)

```{r chicken_supplier}
missed_visitor_ft_in_second_half=pbp_curr_yr %>%
  filter(actionType=="Free Throw",
         #only misses
         str_starts(description,"MISS"),
         #visitor
         location=="v",
         #second half
         period>2,
         #multiple free throws (not and-1s or sole technicals/flagrants)
         str_ends(subType,"2|3"))

missed_visitor_ft_in_second_half %>%
  #free throw groups have same game clock time
  group_by(personId,playerNameI,gameId,clock,period) %>% 
  summarize(n=n()) %>%
  #filter for multiple misses at same time
  filter(n>1) %>%
  group_by(personId,playerNameI) %>%
  summarize(missed_ft_pairs=n()) %>%
  ungroup() %>%
  left_join(.,player_w_gp_percentages %>% 
              select(personId=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId) %>% 
  select(-c(playerNameI,personId)) %>%
  slice_max(missed_ft_pairs,n=5) %>% 
  gt() %>%
  as_raw_html()
```

## The "Rent-Free" Award (presented by Monica Geller)

team that shoots the most technical/flagrant FTAs (credit to Drummallumin for the idea)

```{r rent_free}
flagrant_tech_fts=pbp_curr_yr %>% 
    filter(
      str_detect(subType,"Free Throw Technical|Free Throw Flagrant")
      ) %>%
  group_by(teamTricode) %>%
  summarize(tech_ft=sum(str_detect(subType,"Free Throw Technical")),
            flagrant_ft=sum(str_detect(subType,"Free Throw Flagrant"))) %>%
  mutate(flagrant_plus_tech_fts=tech_ft+flagrant_ft)

flagrant_tech_fts %>% 
  slice_max(flagrant_plus_tech_fts,n=5) %>% gt() %>%
  as_raw_html()
```

## The "It Ain't Over Till the Fat Lady Sings" Award (presented by Kim Kardashian)

most comeback wins where the opposing team had a lead of at least 15 points at some point in the game

```{r largest_leads}
largest_leads=pbp_curr_yr %>%
  filter(!is.na(teamTricode), location %in% c("h", "v")) %>%
  group_by(gameId) %>%
  summarize(
    vis_final = max(scoreAway, na.rm = TRUE),
    home_final = max(scoreHome, na.rm = TRUE),
    largest_home_lead = max(scoreMargin, na.rm = TRUE),
    largest_vis_lead = -min(scoreMargin, na.rm = TRUE),
    home = teamTricode[location == "h"][1],
    visitor = teamTricode[location == "v"][1],
    .groups = "drop"
  ) %>%
  mutate(
    winner_team = if_else(home_final > vis_final, home, visitor),
    loser_team = if_else(winner_team == home, visitor, home),
    largest_winner_lead = if_else(
      winner_team == home, largest_home_lead, largest_vis_lead),
    largest_loser_lead = if_else(
      winner_team == home, largest_vis_lead, largest_home_lead)
  )
```

```{r fat_lady_sings}
largest_leads %>%
  group_by(winner_team) %>%
  summarize(
    num_losses_lead_by_15 = sum(largest_loser_lead >= 15, na.rm = TRUE),
    .groups = "drop"
  ) %>% slice_max(num_losses_lead_by_15,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Snatching Defeat from the Jaws of Victory" Award (presented by the 28-3 Atlanta Falcons)

most losses when having a lead of at least 15 points at some point in the game

```{r snatching_defeat}
largest_leads %>%
  group_by(loser_team) %>%
  summarize(
    num_losses_lead_by_15 = sum(largest_loser_lead >= 15, na.rm = TRUE),
    .groups = "drop"
  ) %>% slice_max(num_losses_lead_by_15,n=5) %>% gt() %>%
  as_raw_html()
```

## The Dikembe Mutombo Memorial "No Fly Zone" Award\*

most blocked dunks as the blocking player

```{r no_fly_zone}
blocked_dunks=inner_join(
  #missed dunks
  pbp_curr_yr %>% filter(shotResult=="Missed",str_detect(subType,"Dunk")),
  #blocks
  pbp_curr_yr %>% filter(str_detect(description,'BLOCK')),
  by=join_by(gameId,clock,period))

#as the blocker
blocked_dunks %>% group_by(personId.y,playerNameI.y) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId.y=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId.y) %>% 
  select(-c(playerNameI.y,personId.y)) %>%
  slice_max(count,n=5) %>% 
  gt() %>% as_raw_html()
```

## The Rejected for Boarding Award (sponsored by United Airlines)\*

most blocked dunks as the dunking player (credit to Legdrop_soup for the idea and asw7412 for the sponsor)

```{r reject_boarding}
#as the blockee
blocked_dunks %>% group_by(personId.x,playerNameI.x) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId.x=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId.x) %>% 
  select(-c(playerNameI.x,personId.x)) %>%
  slice_max(count,n=5) %>% 
  gt() %>% as_raw_html()
```

## The No Time to Taunt Award (presented by Tim Duncan)\*

highest percent of blocks that stayed inbounds & recovered by blocker's team, min 0.5 blocks per game (credit to gibberisle for the idea)

```{r no_time_to_taunt}

blocks_w_next_play=pbp_curr_yr %>% 
  group_by(gameId) %>%
  mutate(next_play=lead(description),
         next_location=lead(location)) %>% ungroup() %>% 
    filter(str_detect(description,'BLOCK')) %>%
  mutate(recovered_by=
           if_else(
             str_detect(next_play,'Rebound',negate = TRUE) & 
              location==next_location,
             "own","other"
           )
  )

player_blk_info=blocks_w_next_play %>% 
  group_by(personId,playerNameI,recovered_by) %>% 
  summarize(blocks=n()) %>% ungroup() %>%
  pivot_wider(names_from=recovered_by,
              values_from=blocks,
              values_fill=0,
              names_glue = "blocks_recovered_by_{recovered_by}") %>%
  mutate(blocks=blocks_recovered_by_own+blocks_recovered_by_other,
         percent_blk_recovered_by_own=blocks_recovered_by_own/blocks) %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId) %>% select(-playerNameI)

average_games_played=team_info %>% 
  summarize(average_GP=mean(TeamGP)) %>% pull(average_GP)

player_blk_info %>% filter(blocks>=average_games_played*0.5) %>% 
  slice_max(percent_blk_recovered_by_own,n=5) %>%
  select(player_name,blocks,own=blocks_recovered_by_own,
         percent_blk_recovered_by_own) %>%
  gt() %>%
  fmt_percent(columns=percent_blk_recovered_by_own) %>%
  as_raw_html()
```

## The "David vs Goliath" Award (presented by Dwyane Wade)

most shots blocked where the blocker is at least 5 inches shorter than the blockee

```{r david_v_goliath}

blocked_shots_w_player_heights=inner_join(
  #missed shots
  pbp_curr_yr %>% filter(shotResult=="Missed"),
  #blocks
  pbp_curr_yr %>% filter(str_detect(description,'BLOCK')),
  by=join_by(gameId,clock,period)) %>%
  left_join(.,player_w_gp_percentages %>% 
              select(PLAYER_ID,
                     blocked_player_height=PLAYER_HEIGHT_INCHES),
            by=join_by(personId.x==PLAYER_ID)) %>%
  left_join(.,player_w_gp_percentages %>% 
              select(PLAYER_ID,
                     blocking_player_height=PLAYER_HEIGHT_INCHES),
            by=join_by(personId.y==PLAYER_ID)) %>% 
  mutate(height_diff=blocked_player_height-blocking_player_height)

blocked_shots_w_player_heights %>% filter(height_diff>=5) %>%
  group_by(personId.y,playerNameI.y) %>% summarize(count=n()) %>%
  ungroup() %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId.y=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId.y) %>% 
  select(-c(playerNameI.y,personId.y)) %>%
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Call Game" Award (presented by Paul "No OT Tonight" George)

most game winning points (defined as the first points that eclipsed the losing team's total) (credit to Necessary_Career_253 for the idea & Clownp3nis for the presenter)

```{r game_winners}
game_winning_shots=pbp_curr_yr %>% group_by(gameId) %>% 
  mutate(vis_final=max(scoreAway),
            home_final=max(scoreHome)) %>% 
  ungroup() %>%
  mutate(winner=if_else(home_final>vis_final,"home","visitor"),
         loser_points=if_else(home_final>vis_final,
                              vis_final,home_final)) %>%
  filter((winner=="home" & scoreHome>loser_points)|
           (winner=="visitor" & scoreAway>loser_points)) %>%
  group_by(gameId) %>% slice_min(actionId) %>% ungroup() %>%
  mutate(shotValue=case_when(
    actionType=="Free Throw"~1,
    TRUE~shotValue
  ))

game_win_shot_summary=game_winning_shots %>%
  group_by(personId,playerNameI) %>% 
  summarize(game_winning_3=sum(shotValue==3),
            game_winning_2=sum(shotValue==2),
            game_winning_ft=sum(shotValue==1),
            game_winning_points=sum(shotValue)) %>% 
  ungroup() %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId) %>% 
  select(-c(playerNameI))

write_csv(game_win_shot_summary,
          "Output Data/Game-Winning Shots.csv")

game_win_shot_summary %>% 
  slice_max(game_winning_points,n=5) %>%
  select(-personId) %>%
  gt() %>%
  cols_label(game_winning_3="GW3",game_winning_2="GW2",
             game_winning_ft="GWFT",game_winning_points="GWPTS") %>%
  as_raw_html()
```

## The Bowling Ball Award (sponsored by [Pete Weber](#0), presented by Glen "Big Baby" Davis)\*

most charges committed (credit to Kdog122025 for the idea)

```{r bowling_ball}
violations=pbp_curr_yr %>% 
  filter(
    str_detect(
      description,
      "Charge|Goaltending|Kick|Traveling|3 Second")) %>%
  left_join(.,player_w_gp_percentages %>% 
              select(personId=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId) %>%
  select(-c(playerNameI))
  
violations %>% 
  filter(str_detect(description,"Charge")) %>% 
  group_by(player_name) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## "The Good Ol' Hockey Game, is the Best Game You Can Name" Award (presented by Dominik Hasek)\*

most goaltends committed (credit to Kdog122025 for the idea)

```{r goaltends}
violations %>% 
  filter(str_detect(description,"Goaltending")) %>% 
  group_by(player_name) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## "The Thing about Arsenal Is They Always Try to Walk It In" Award (presented by MLS Commissioner Don Garber)\*

most kicked ball violations

```{r kicked_balls}
violations %>% 
  filter(str_detect(description,"Kick")) %>% 
  group_by(player_name) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Pack Your Bags" Award (Sponsored by Emirates, the global airline partner of the NBA)

Player with the most traveling calls

```{r travelling}
violations %>% 
  filter(str_detect(description,"Traveling")) %>% 
  group_by(player_name) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Counting to 3" Award (Sponsored by the National Parks System, presented by Sesame Street's Count von Count)

Player with the most illegal defense/defensive 3-second calls (credit to PsychoM & MrBuckBuck for the idea)

```{r def_3_seconds}
violations %>% 
  filter(str_detect(description,"3 Second")) %>% 
  group_by(player_name) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "Time is An Illusion" Award (sponsored by Salvador Dali)

team with the most 24-second shotclock, 8-second backcourt and 5-second inbound violations (credit to Necessary_Career_253 for the idea)

```{r time_is_an_illusion}
team_violations=pbp_curr_yr %>% filter(
  actionType=="Turnover",
  str_detect(subType,"Shot Clock|5 Second|8 Second")
) %>%
  mutate(team_name=str_to_title(word(description,1,
                                     sep=" Turnover:")))

team_violations %>%
  group_by(team_name) %>%
  summarize(shot_clock=sum(str_detect(subType,"Shot Clock")),
            inbound=sum(str_detect(subType,"5 Second")),
            halfcourt=sum(str_detect(subType,"8 Second"))
            ) %>%
  mutate(tot_violations=shot_clock+inbound+halfcourt) %>%
  slice_max(tot_violations,n=5) %>%
  gt() %>% as_raw_html()
```

## The Zaza Pachulia All-Ball Award\*

most 3-pt shooting fouls committed (credit to watchingsongsDL, kingcobweb & An-Indian-In-The-NBA for the idea, and sunnysideoutside for the name)

```{r 3_pt_fouls_commit}

fouls_on_threes=inner_join(
  #shooting fouls
  pbp_curr_yr %>% filter(str_detect(description,"S.FOUL")),
  pbp_curr_yr %>%
    filter(
      #missed threes
      str_detect(description,"Free Throw 3 of 3")|
        #made threes
        (str_detect(description,"3PT") & shotResult=="Made")),
  by=c('gameId','period','clock')) %>% ungroup()

fouls_on_threes %>% group_by(personId.x,playerNameI.x) %>% 
  summarize(count=n()) %>% ungroup() %>% 
  left_join(.,player_w_gp_percentages %>% 
              select(personId.x=PLAYER_ID,player_name=PLAYER_NAME)) %>%
  relocate(player_name,.after=personId.x) %>%
  select(-c(playerNameI.x,personId.x)) %>%
  slice_max(count,n=5) %>% gt() %>%
  as_raw_html()
```

## The "I'll Have It to Go" Award (sponsored by DoorDash)

coach with lowest timeout utilization (credit to xfinityhomeboy, Ill_Ad3517 & s-sea (who also came up with the name))

```{r most_timeouts_left_over}
num_timeouts_available=pbp_curr_yr %>% 
  filter(!is.na(teamTricode)) %>% 
  group_by(gameId,period,teamTricode) %>% slice(1) %>% 
  select(gameId,period,teamTricode) %>% ungroup() %>%
  group_by(team=teamTricode) %>% 
  summarize(games=sum(period<=4)/4,
            overtimes=sum(period>4),
            num_timeouts=7*games+2*overtimes) %>% 
  left_join(.,team_info %>% 
              select(TeamName,TEAM_ABBREVIATION),
            by=join_by(team==TEAM_ABBREVIATION))

num_timeouts_used=pbp_curr_yr %>% 
  filter(str_detect(description,"Timeout")) %>% 
  filter(str_detect(description,"Excess",negate=TRUE)) %>%
  select(gameId,period,description) %>%
  mutate(team_name=str_to_title(word(description,1,
                                     sep=" Timeout:"))) %>%
  group_by(team_name) %>% summarise(num_timeouts_used=n())

timeouts_used_percent=
  left_join(num_timeouts_available,num_timeouts_used,
            by=join_by(TeamName==team_name)) %>% 
  mutate(timeout_use_percent=num_timeouts_used/num_timeouts) %>% 
  arrange(desc(timeout_use_percent))

timeouts_used_percent %>% 
  select(team,num_timeouts,
         num_timeouts_used,timeout_use_percent) %>%
  slice_min(timeout_use_percent,n=5) %>%
  gt() %>%
  fmt_percent(columns=timeout_use_percent) %>%
  as_raw_html()
```

## The "Hotheaded" Award (presented by Don Nelson)

most technicals plus ejections for non-players (credit to livejamie for the idea)

```{r hotheaded}
coach_ejection_techs=pbp_curr_yr %>% filter(
  teamId==0,
  actionType =="Ejection"|
    str_detect(description,"T.FOUL|DOUBLE.TECHNICAL.FOUL")) %>%
    mutate(coach=if_else(
      actionType=="Ejection",
      str_to_title(word(description,1,sep=" Ejection:")),
      str_to_title(word(description,1,sep=" Foul:"))
    )) 

coach_ejection_techs %>%
  group_by(coach) %>%
  summarize(ejections=sum(actionType=="Ejection"),
            technicals=sum(actionType=="Foul")) %>%
  mutate(techs_plus_eject=technicals+ejections) %>%
  slice_max(techs_plus_eject,n=5) %>% gt() %>%
  as_raw_html()
```

\newpage

\KOMAoptions{usegeometry,paper=landscape,pagesize}

\recalctypearea

\newgeometry{left=0.5in, right=0.5in, top=0.5in, bottom=0.5in}

## The "Welcome to the League, Rook" Award (sponsored by Orville Redenbacher)

most combined times stealing from a rookie, blocking a rookie and getting free throws after being fouled by a rookie

```{r events_against_rookies_work}

#get rookies
rookies=nba_commonallplayers(is_only_current_season=1)$CommonAllPlayers %>%
  type_convert() %>%
  filter(FROM_YEAR==2025)

events_by_rookies=inner_join(pbp_curr_yr,rookies %>% select(PERSON_ID),
             by=join_by(personId==PERSON_ID))

#fouls and free throws have same timing, hence inner join
free_throws_on_fouls_by_rookies=
  inner_join(
    #want to measure free throw trips, not total free throws
    pbp_curr_yr %>% filter(str_detect(description,"1 of")),
    #get fouls only committed by rookies
    events_by_rookies %>% filter(actionType=="Foul"),
    by=join_by(gameId,clock,period)) %>%
  #remove instances of technicals by same team player
  filter(teamTricode.x != teamTricode.y)

fouls_on_free_throws_by_rookies=
  inner_join(pbp_curr_yr %>% filter(actionType=="Foul"),
             #get free throw trips by rookies
             events_by_rookies %>% filter(str_detect(description,"1 of")),
             by=join_by(gameId,clock,period)) %>%
  #remove instances of technicals by same team player
  filter(teamTricode.x != teamTricode.y)

rookie_shots_blocked=
  inner_join(pbp_curr_yr %>% filter(str_detect(description,"BLOCK")),
             events_by_rookies %>% filter(shotResult=="Missed"),
             by=join_by(gameId,clock,period))

shots_blocked_by_rookies=
  inner_join(pbp_curr_yr %>% filter(shotResult=="Missed"),
             events_by_rookies %>% filter(str_detect(description,"BLOCK")),
             by=join_by(gameId,clock,period))

steals_on_turnovers_by_rookies =
  inner_join(pbp_curr_yr %>% filter(str_detect(description,"STEAL")),
             #steals are either bad pass or lost ball
             events_by_rookies %>% 
               filter(actionType=="Turnover",
                      subType %in% c("Bad Pass","Lost Ball")),
             by=join_by(gameId,clock,period))

turnovers_on_steals_by_rookies =
  inner_join(
    #steals are either bad pass or lost ball
    pbp_curr_yr %>% 
      filter(actionType=="Turnover",subType %in% c("Bad Pass","Lost Ball")),
    events_by_rookies %>% filter(str_detect(description,"STEAL")),
    by=join_by(gameId,clock,period))

events_against_rookies=
  bind_rows(free_throws_on_fouls_by_rookies,
            rookie_shots_blocked,
            steals_on_turnovers_by_rookies,
            fouls_on_free_throws_by_rookies,
            shots_blocked_by_rookies,
            turnovers_on_steals_by_rookies)

rm(free_throws_on_fouls_by_rookies,
            rookie_shots_blocked,
            steals_on_turnovers_by_rookies,
            fouls_on_free_throws_by_rookies,
            shots_blocked_by_rookies,
            turnovers_on_steals_by_rookies)
  
events_against_rookies_summ=
  events_against_rookies %>%
  group_by(personId.x,playerNameI.x) %>%
  summarize(num_times_stolen_from_rookie=
              sum(str_detect(description.x,"STEAL")),
            num_times_stolen_by_rookie=
              sum(str_detect(description.y,"STEAL")),
            num_times_fouled_by_rookie=
              sum(str_detect(description.x,"1 of")),
            num_times_fouled_rookie=
              sum(str_detect(description.y,"1 of")),
            num_times_blocked_rookie=
              sum(str_detect(description.x,"BLOCK")),
            num_times_blocked_by_rookie=
              sum(str_detect(description.y,"BLOCK")),
            net_steals_rookies=
              num_times_stolen_from_rookie-num_times_stolen_by_rookie,
            net_fouls_rookies=
              num_times_fouled_by_rookie-num_times_fouled_rookie,
            net_blocks_rookies=
              num_times_blocked_rookie-num_times_blocked_by_rookie,
            positive_events_against_rookies=
              num_times_stolen_from_rookie+num_times_fouled_by_rookie+
              num_times_blocked_rookie,
            negative_events_against_rookies=
              num_times_stolen_by_rookie+num_times_fouled_rookie+
              num_times_blocked_by_rookie,
            net_events_against_rookies=
              positive_events_against_rookies-negative_events_against_rookies
            ) %>% 
  ungroup() %>%
  left_join(.,
            player_w_gp_percentages %>% 
              select(personId.x=PLAYER_ID,PLAYER_NAME,G_PERCENT,MIN)) %>%
  relocate(PLAYER_NAME) %>% select(-c(personId.x,playerNameI.x)) %>%
  mutate(
    across(
      positive_events_against_rookies:net_events_against_rookies,
      ~./MIN*36,
      .names="{.col}_per_36"))
```

```{r welcome_to_the_league_rook}
#| tbl-pos: T
#| layout-ncol: 1

events_against_rookies_summ %>% 
  filter(G_PERCENT>=0.5,MIN>=300) %>%
  select(player=PLAYER_NAME,min=MIN,
         num_times_stolen_from_rookie,num_times_fouled_by_rookie,
         num_times_blocked_rookie,num_times_stolen_by_rookie,
         num_times_fouled_rookie,num_times_blocked_by_rookie,
         positive_events_against_rookies_per_36:
           net_events_against_rookies_per_36) %>%
  slice_max(net_events_against_rookies_per_36,n=5) %>%
  gt() %>%
  tab_spanner(label="Positive Events",3:5) %>%
  tab_spanner(label="Negative Events",6:8) %>%
  tab_spanner(label="Per 36",9:11) %>%
  cols_label(num_times_stolen_from_rookie="Steals Against Rookie",
             num_times_stolen_by_rookie="Steals By Rookie",
             num_times_fouled_by_rookie="Fouled By Rookie",
             num_times_fouled_rookie="Foul Against Rookie",
             num_times_blocked_rookie="Blocks Against Rookie",
             num_times_blocked_by_rookie="Blocks by Rookie",
             positive_events_against_rookies_per_36="Positive Events",
             negative_events_against_rookies_per_36="Negative Events",
             net_events_against_rookies_per_36="Net Events") %>%
  fmt_number(ends_with("per_36"),decimals=4) %>%
  cols_width(player~px(150),
             everything()~px(70))
```

## The "The Future is Now, Old Man" Award (sponsored by the National Basketball Retired Players Association)

most combined times getting blocked by a rookie, getting the ball stolen by a rookie and fouling a rookie who takes free throws

```{r future_is_now_old_man}
#| tbl-pos: T
#| layout-ncol: 1

events_against_rookies_summ %>% 
  filter(G_PERCENT>=0.5,MIN>=300) %>%
  select(player=PLAYER_NAME,min=MIN,
         num_times_stolen_from_rookie,num_times_fouled_by_rookie,
         num_times_blocked_rookie,num_times_stolen_by_rookie,
         num_times_fouled_rookie,num_times_blocked_by_rookie,
         positive_events_against_rookies_per_36:
           net_events_against_rookies_per_36) %>%
  slice_min(net_events_against_rookies_per_36,n=5) %>%
  gt() %>%
  tab_spanner(label="Positive Events",3:5) %>%
  tab_spanner(label="Negative Events",6:8) %>%
  tab_spanner(label="Per 36",9:11) %>%
  cols_label(num_times_stolen_from_rookie="Steals Against Rookie",
             num_times_stolen_by_rookie="Steals By Rookie",
             num_times_fouled_by_rookie="Fouled By Rookie",
             num_times_fouled_rookie="Foul Against Rookie",
             num_times_blocked_rookie="Blocks Against Rookie",
             num_times_blocked_by_rookie="Blocks by Rookie",
             positive_events_against_rookies_per_36="Positive Events",
             negative_events_against_rookies_per_36="Negative Events",
             net_events_against_rookies_per_36="Net Events") %>%
  fmt_number(ends_with("per_36"),decimals=4) %>%
  cols_width(player~px(150),
             everything()~px(70))
```
